---
title: "Primary Modeling: County Economic Recovery, 2008 vs. COVID"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    df-print: kable
execute:
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(estimatr)   # lm_robust, HC1 SEs
library(glmnet)
library(randomForest)

set.seed(20260729)

county <- read_csv("spine_data.csv", show_col_types = FALSE)
dir.create("raw_data", showWarnings = FALSE)
```

Read-only on `spine_data.csv`. Outcome is `log(recovery_ratio)` in both
crises, no exceptions. Identical predictor set, control set, and transform
for both crises -- the only thing that varies between the two calls below is
which year's columns feed the same function.

## Data prep (one function, called twice)

```{r crisis-cols}
# The ONE crisis-specific thing in this whole document: which source column
# maps to which generic name, per crisis.
crisis_cols <- list(
  "2008" = c(outcome = "recovery_ratio_2008", flag_tiny = "flag_tiny_base_2008",
             proprietor_share = "proprietor_share_2007", migration_rate = "net_migration_rate_2007",
             unemployment_rate = "pre_unemployment_2007", working_age_share = "working_age_share_2007",
             income = "median_hh_income_2007", poverty_rate = "poverty_rate_2007",
             population = "total_population_2007", density = "population_density_2007",
             rucc = "rucc_2003_2007", pop_trend = "pop_trend_2007"),
  "covid" = c(outcome = "recovery_ratio_covid", flag_tiny = "flag_tiny_base_covid",
              proprietor_share = "proprietor_share_2019", migration_rate = "net_migration_rate_2019",
              unemployment_rate = "pre_unemployment_2019", working_age_share = "working_age_share_2019",
              income = "median_hh_income_2019", poverty_rate = "poverty_rate_2019",
              population = "total_population_2019", density = "population_density_2019",
              rucc = "rucc_2013_2019", pop_trend = "pop_trend_2019")
)

# Studied predictors get standardized WITHIN this crisis's complete-case
# sample; controls are used on their stated scale (raw or logged) and never
# standardized -- they're not being interpreted, only held constant.
prepare_crisis_data <- function(raw, crisis) {
  cols <- crisis_cols[[crisis]]
  raw %>%
    transmute(
      fips, county_name,
      outcome_raw = .data[[cols[["outcome"]]]],
      flag_tiny_base = .data[[cols[["flag_tiny"]]]],
      proprietor_share = .data[[cols[["proprietor_share"]]]],
      migration_rate = .data[[cols[["migration_rate"]]]],
      unemployment_rate = .data[[cols[["unemployment_rate"]]]],
      working_age_share = .data[[cols[["working_age_share"]]]],
      income = .data[[cols[["income"]]]],
      poverty_rate = .data[[cols[["poverty_rate"]]]],
      population = .data[[cols[["population"]]]],
      density = .data[[cols[["density"]]]],
      rucc_raw = .data[[cols[["rucc"]]]],
      pop_trend = .data[[cols[["pop_trend"]]]]
    ) %>%
    filter(if_all(c(outcome_raw, proprietor_share, migration_rate, unemployment_rate, working_age_share,
                     income, poverty_rate, population, density, rucc_raw, pop_trend), ~ !is.na(.))) %>%
    mutate(
      log_outcome = log(outcome_raw),
      log_population = log(population),
      log_density = log(density),
      rucc = relevel(factor(rucc_raw), ref = "1"),  # unordered factor(), level 1 = reference
      proprietor_share_z = as.numeric(scale(proprietor_share)),
      migration_rate_z = as.numeric(scale(migration_rate)),
      unemployment_rate_z = as.numeric(scale(unemployment_rate)),
      working_age_share_z = as.numeric(scale(working_age_share))
    )
}

studied_z <- c("proprietor_share_z", "migration_rate_z", "unemployment_rate_z", "working_age_share_z")
controls  <- c("income", "poverty_rate", "log_population", "log_density", "rucc", "pop_trend")
controls_no_wealth <- setdiff(controls, c("income", "poverty_rate"))
```

```{r run-crisis-models}
# Everything below is written once and driven by `crisis`; no 2008 block
# copy-pasted into a COVID block.
run_crisis_models <- function(raw, crisis) {
  df <- prepare_crisis_data(raw, crisis)

  ## 1. Pooled OLS, HC1 robust SEs (diagnostics showed migration variance
  ## fanning out in small-population counties -- default SEs would be wrong)
  ols_formula <- as.formula(paste("log_outcome ~", paste(c(studied_z, controls), collapse = " + ")))
  ols <- lm_robust(ols_formula, data = df, se_type = "HC1")

  ## 2. LASSO -- controls (incl. all 8 RUCC dummies) unpenalized, studied
  ## predictors penalized. penalty.factor = 0 keeps a control in the model
  ## regardless of magnitude, so "holding X constant" stays true.
  X <- model.matrix(as.formula(paste("~", paste(c(studied_z, controls), collapse = " + "))), data = df)[, -1]
  y <- df$log_outcome
  penalty <- ifelse(colnames(X) %in% studied_z, 1, 0)
  set.seed(20260729)
  cvfit <- cv.glmnet(X, y, alpha = 1, nfolds = 10, standardize = TRUE, penalty.factor = penalty)

  ## 3a. RF WITH controls incl. income/poverty -- prediction accuracy only.
  rf_formula_a <- as.formula(paste("log_outcome ~", paste(c(studied_z, controls), collapse = " + ")))
  set.seed(20260729)
  rf_a <- randomForest(rf_formula_a, data = df, ntree = 500, importance = FALSE)

  ## 3b. RF WITHOUT income/poverty -- so the 4 studied predictors compete
  ## against each other and the remaining controls, not against wealth.
  rf_formula_b <- as.formula(paste("log_outcome ~", paste(c(studied_z, controls_no_wealth), collapse = " + ")))
  set.seed(20260729)
  rf_b <- randomForest(rf_formula_b, data = df, ntree = 500, importance = TRUE)

  list(crisis = crisis, df = df, n = nrow(df), ols = ols, cvfit = cvfit,
       coef_min = as.matrix(coef(cvfit, s = "lambda.min"))[, 1],
       coef_1se = as.matrix(coef(cvfit, s = "lambda.1se"))[, 1],
       rf_a = rf_a, rf_b = rf_b)
}
```

```{r fit-or-load}
# Cache fitted objects -- LASSO's CV and 500-tree RF x2 x2 crises aren't
# instant; no need to refit on every render.
cache_path <- "raw_data/modeling_andrew_results.rds"
if (file.exists(cache_path)) {
  results <- readRDS(cache_path)
} else {
  results <- list(
    "2008" = run_crisis_models(county, "2008"),
    "covid" = run_crisis_models(county, "covid")
  )
  saveRDS(results, cache_path)
}
res_2008 <- results[["2008"]]
res_covid <- results[["covid"]]

tibble(crisis = c("2008", "covid"), n = c(res_2008$n, res_covid$n))
```

## 1. Pooled OLS (HC1 robust SEs)

```{r ols-studied}
ols_studied_table <- function(res) {
  s <- summary(res$ols)$coefficients[studied_z, ]
  as_tibble(s, rownames = "term") %>%
    rename(estimate = Estimate, robust_se = `Std. Error`, t = `t value`,
           p = `Pr(>|t|)`, ci_low = `CI Lower`, ci_high = `CI Upper`) %>%
    select(term, estimate, robust_se, t, p, ci_low, ci_high)
}
bind_rows("2008" = ols_studied_table(res_2008), "covid" = ols_studied_table(res_covid), .id = "crisis")
```

```{r ols-fit-stats}
tibble(
  crisis = c("2008", "covid"),
  adj_r_squared = c(res_2008$ols$adj.r.squared, res_covid$ols$adj.r.squared),
  residual_sd = c(sqrt(res_2008$ols$res_var), sqrt(res_covid$ols$res_var))
)
```

### Appendix: control coefficients (not interpreted, reported for completeness)

```{r ols-controls-appendix}
ols_controls_table <- function(res) {
  s <- summary(res$ols)$coefficients
  as_tibble(s, rownames = "term") %>%
    filter(!term %in% studied_z, term != "(Intercept)") %>%
    rename(estimate = Estimate, robust_se = `Std. Error`, p = `Pr(>|t|)`) %>%
    select(term, estimate, robust_se, p)
}
bind_rows("2008" = ols_controls_table(res_2008), "covid" = ols_controls_table(res_covid), .id = "crisis")
```

## 2. LASSO (10-fold CV, controls + RUCC dummies unpenalized)

```{r lasso-table}
lasso_table <- function(res) {
  tibble(term = names(res$coef_min), lambda_min = res$coef_min, lambda_1se = res$coef_1se) %>%
    filter(term %in% studied_z)
}
lasso_results <- bind_rows("2008" = lasso_table(res_2008), "covid" = lasso_table(res_covid), .id = "crisis")
lasso_results %>% mutate(survives_1se = lambda_1se != 0)
```

```{r lasso-full-coefs}
# full coefficient vectors (controls included) at lambda.1se, for reference
tibble(term = names(res_2008$coef_1se), coef_2008 = res_2008$coef_1se) %>%
  left_join(tibble(term = names(res_covid$coef_1se), coef_covid = res_covid$coef_1se), by = "term")
```

## 3. Random forest (500 trees), run twice per crisis

### 3a. WITH controls (income/poverty included) -- prediction accuracy only

```{r rf-a-fit}
rf_a_stats <- tibble(
  crisis = c("2008", "covid"),
  oob_r_squared = c(res_2008$rf_a$rsq[500], res_covid$rf_a$rsq[500]),
  oob_rmse = c(sqrt(res_2008$rf_a$mse[500]), sqrt(res_covid$rf_a$mse[500]))
)
rf_a_stats
```

### 3b. WITHOUT income/poverty -- permutation importance only

```{r rf-b-importance}
rf_b_importance <- function(res) {
  imp <- importance(res$rf_b, type = 1)
  tibble(term = rownames(imp), pct_inc_mse = imp[, 1]) %>% arrange(desc(pct_inc_mse))
}
bind_rows("2008" = rf_b_importance(res_2008), "covid" = rf_b_importance(res_covid), .id = "crisis")
```

## 4. Comparison table

```{r comparison-table}
rf_b_rank <- function(res) {
  imp <- importance(res$rf_b, type = 1)
  tibble(term = rownames(imp), pct_inc_mse = imp[, 1]) %>%
    mutate(rank = rank(-pct_inc_mse)) %>%
    filter(term %in% studied_z)
}

build_comparison <- function(res, crisis_label) {
  ols_t <- ols_studied_table(res) %>% select(term, ols_estimate = estimate, ols_robust_se = robust_se)
  lasso_t <- lasso_table(res) %>% mutate(survives_1se = lambda_1se != 0) %>% select(term, survives_1se)
  rf_t <- rf_b_rank(res) %>% select(term, rf_importance_rank = rank)
  ols_t %>% left_join(lasso_t, by = "term") %>% left_join(rf_t, by = "term") %>% mutate(crisis = crisis_label)
}

comparison <- bind_rows(build_comparison(res_2008, "2008"), build_comparison(res_covid, "covid")) %>%
  select(term, crisis, ols_estimate, ols_robust_se, survives_1se, rf_importance_rank)
comparison
```

```{r comparison-signs}
sign_flip <- comparison %>%
  select(term, crisis, ols_estimate) %>%
  pivot_wider(names_from = crisis, values_from = ols_estimate) %>%
  mutate(sign_flip = sign(`2008`) != sign(covid))
sign_flip
```

## Robustness checks

These are checks on the primary result, not a replacement for it.

```{r robustness-tinybase}
# union of flag_tiny_base across BOTH crises, computed on the full table --
# same counties dropped from both samples, so the 2008-vs-COVID comparison
# doesn't break by comparing different populations.
tiny_union_fips <- county %>% filter(flag_tiny_base_2008 | flag_tiny_base_covid) %>% pull(fips)

run_ols_variant <- function(df, formula_rhs = paste(c(studied_z, controls), collapse = " + ")) {
  lm_robust(as.formula(paste("log_outcome ~", formula_rhs)), data = df, se_type = "HC1")
}

ols_notiny_2008 <- run_ols_variant(res_2008$df %>% filter(!fips %in% tiny_union_fips))
ols_notiny_covid <- run_ols_variant(res_covid$df %>% filter(!fips %in% tiny_union_fips))

cat("Tiny-base union size:", length(tiny_union_fips), "counties, excluded from both crises\n")
cat("n after exclusion -- 2008:", nrow(res_2008$df %>% filter(!fips %in% tiny_union_fips)),
    " covid:", nrow(res_covid$df %>% filter(!fips %in% tiny_union_fips)), "\n")

bind_rows(
  "2008" = as_tibble(summary(ols_notiny_2008)$coefficients[studied_z, c("Estimate", "Pr(>|t|)")], rownames = "term"),
  "covid" = as_tibble(summary(ols_notiny_covid)$coefficients[studied_z, c("Estimate", "Pr(>|t|)")], rownames = "term"),
  .id = "crisis"
)
```

```{r robustness-winsorized}
# outcome (already log-scale, per the project's stated transform) winsorized
# at its own 1st/99th percentile, per crisis
winsorize <- function(x, p = 0.01) {
  bounds <- quantile(x, c(p, 1 - p))
  pmin(pmax(x, bounds[1]), bounds[2])
}

ols_wins_2008 <- run_ols_variant(res_2008$df %>% mutate(log_outcome = winsorize(log_outcome)))
ols_wins_covid <- run_ols_variant(res_covid$df %>% mutate(log_outcome = winsorize(log_outcome)))

bind_rows(
  "2008" = as_tibble(summary(ols_wins_2008)$coefficients[studied_z, c("Estimate", "Pr(>|t|)")], rownames = "term"),
  "covid" = as_tibble(summary(ols_wins_covid)$coefficients[studied_z, c("Estimate", "Pr(>|t|)")], rownames = "term"),
  .id = "crisis"
)
```

```{r robustness-comparison}
# side-by-side against the primary OLS estimates, for an at-a-glance sign/magnitude check
primary <- comparison %>% filter(crisis %in% c("2008","covid")) %>%
  select(term, crisis, primary_estimate = ols_estimate) %>% distinct()

robustness_compare <- bind_rows(
  as_tibble(summary(ols_notiny_2008)$coefficients[studied_z, "Estimate", drop=FALSE], rownames="term") %>%
    mutate(crisis = "2008", variant = "tiny_base_union_excluded") %>% rename(estimate = Estimate),
  as_tibble(summary(ols_notiny_covid)$coefficients[studied_z, "Estimate", drop=FALSE], rownames="term") %>%
    mutate(crisis = "covid", variant = "tiny_base_union_excluded") %>% rename(estimate = Estimate),
  as_tibble(summary(ols_wins_2008)$coefficients[studied_z, "Estimate", drop=FALSE], rownames="term") %>%
    mutate(crisis = "2008", variant = "winsorized_outcome") %>% rename(estimate = Estimate),
  as_tibble(summary(ols_wins_covid)$coefficients[studied_z, "Estimate", drop=FALSE], rownames="term") %>%
    mutate(crisis = "covid", variant = "winsorized_outcome") %>% rename(estimate = Estimate)
) %>%
  left_join(primary, by = c("term", "crisis")) %>%
  mutate(same_sign = sign(estimate) == sign(primary_estimate)) %>%
  select(term, crisis, variant, primary_estimate, estimate, same_sign)

robustness_compare
```

## Section A: Wealth-band split

Terciles of pre-crisis income, computed **separately per crisis year** and
**on that crisis's own modeling sample** (not reusing the full-spine
`wealth_band_*` columns, which were cut on a different, larger population and
would produce lopsided bands here) -- same OLS spec, run within each band,
income and poverty still included as controls inside every band. 6 models
total.

```{r wealthband-setup}
add_income_band <- function(df) {
  df %>% mutate(income_band = cut(income, breaks = quantile(income, c(0, 1/3, 2/3, 1)),
                                   include.lowest = TRUE, labels = c("lower", "middle", "higher")))
}
df_2008_banded <- add_income_band(res_2008$df)
df_covid_banded <- add_income_band(res_covid$df)

bind_rows(
  "2008" = count(df_2008_banded, income_band),
  "covid" = count(df_covid_banded, income_band),
  .id = "crisis"
)
```

```{r wealthband-fit}
# same run_ols_variant() helper as the robustness section, just subset by band
run_band_ols <- function(df, band) {
  run_ols_variant(df %>% filter(income_band == band))
}

bands <- c("lower", "middle", "higher")
wealthband_cache <- "raw_data/modeling_andrew_wealthband_fits.rds"
if (file.exists(wealthband_cache)) {
  band_fits <- readRDS(wealthband_cache)
} else {
  band_fits <- list(
    "2008" = setNames(lapply(bands, function(b) run_band_ols(df_2008_banded, b)), bands),
    "covid" = setNames(lapply(bands, function(b) run_band_ols(df_covid_banded, b)), bands)
  )
  saveRDS(band_fits, wealthband_cache)
}

band_table <- imap_dfr(band_fits, function(crisis_fits, crisis) {
  imap_dfr(crisis_fits, function(fit, band) {
    s <- summary(fit)$coefficients[studied_z, ]
    as_tibble(s, rownames = "term") %>%
      rename(estimate = Estimate, robust_se = `Std. Error`, p = `Pr(>|t|)`) %>%
      mutate(crisis = crisis, band = band)
  })
})

band_table %>%
  select(term, crisis, band, estimate, robust_se, p) %>%
  arrange(term, crisis, factor(band, levels = bands))
```

```{r wealthband-n}
tibble(
  crisis = rep(c("2008", "covid"), each = 3), band = rep(bands, 2),
  n = c(sapply(band_fits[["2008"]], function(f) f$nobs), sapply(band_fits[["covid"]], function(f) f$nobs))
)
```

### Flags

```{r wealthband-sign-flip}
# sign flip across bands, WITHIN the same crisis
sign_by_band <- band_table %>%
  select(term, crisis, band, estimate) %>%
  pivot_wider(names_from = band, values_from = estimate) %>%
  mutate(sign_flip_within_crisis = !(sign(lower) == sign(middle) & sign(middle) == sign(higher)))
sign_by_band
```

```{r wealthband-migration-2008}
# is migration_rate in 2008 (null in the pooled model, p = .75) non-null in ANY band?
band_table %>% filter(term == "migration_rate_z", crisis == "2008") %>% select(band, estimate, p)
```

```{r wealthband-pooled-vs-bands}
# pooled coefficient vs. the 3 band-level coefficients, side by side
pooled <- comparison %>% select(term, crisis, pooled_estimate = ols_estimate) %>% distinct()
band_table %>%
  select(term, crisis, band, band_estimate = estimate, p) %>%
  left_join(pooled, by = c("term", "crisis")) %>%
  arrange(term, crisis, factor(band, levels = bands))
```

## Section B: Placebo -- calm period (2014 to 2017)

Same 4 studied predictors, same 6 controls, same log transform, same OLS/RF
methods as the primary models. Outcome: `log(employment_2017 /
employment_2014)`. Predictors and controls measured 2013 -- strictly before
the window.

**Data-availability notes, since this window wasn't part of the original
pipeline (which only ever pulled 2007/2019 vintages):**

- **2017 county employment isn't in `raw_data/01_bls_panel.rds`** -- the
  original BLS pull deliberately skipped 2017 (`years_needed` was
  `2006:2016, 2018:2025`, since neither crisis window needed it). BLS's own
  site returns HTTP 403 to every automated request from this environment
  (documented in `pipeline/01_bls_employment.R`), so `laucnty17.xlsx` was
  retrieved from the Internet Archive's Wayback Machine
  (`web.archive.org`, snapshot of the same file BLS originally published) --
  a public mirror of the same government file, not a workaround of any
  access control. That file also uses a different (older, multi-row-wrapped)
  header layout than the 2007/2019 files, so it needed its own parser
  rather than reusing `01_bls_employment.R`'s header-detection logic.
- **All 2013 predictor/control values are new**, since the pipeline never
  pulled a 2013 vintage of anything. Reused already-cached raw files where
  possible: BEA `CAEMP25N` bulk file (2001-2022, already has a 2013 column),
  USDA RUCC 2013 vintage (already pulled for the COVID side), Census PEP
  2010s population file (already has `POPESTIMATE2010`/`POPESTIMATE2013`),
  Census Gazetteer land area, and the already-cached per-state
  `cc-est2019-agesex-*.csv` files (`YEAR` code 6 = 2013, same coding scheme
  documented in `pipeline/19_age_structure.R`). Two fresh pulls were needed:
  SAIPE 2013 (income/poverty) and IRS SOI's 2012-2013 county migration pair
  -- both plain API/CSV pulls, no access issues.
- **`pop_trend_2013` uses a 3-year window (2010-2013), not the original
  7-year window.** A 7-year window ending in 2013 (e.g. 2006-2013) would
  cross the Census 2000s-intercensal/2010s-vintage boundary, mixing two
  different population-estimate methodologies in one CAGR calculation --
  the original 2007/2019 windows never had to do this (each stayed inside
  one vintage). 2010-2013 stays inside the single 2010s vintage file
  instead, at the cost of a shorter window.

```{r placebo-assemble}
placebo_cache <- "raw_data/placebo_2013_pieces.rds"
valid_state_fips <- c("01","02","04","05","06","08","09","10","11","12","13","15","16","17",
                       "18","19","20","21","22","23","24","25","26","27","28","29","30","31",
                       "32","33","34","35","36","37","38","39","40","41","42","44","45","46",
                       "47","48","49","50","51","53","54","55","56")

if (file.exists(placebo_cache)) {
  pieces <- readRDS(placebo_cache)
} else {
  ua <- "Mozilla/5.0 (Macintosh; Intel Mac OS X 10_15_7) AppleWebKit/537.36"
  census_key <- Sys.getenv("CENSUS_API_KEY")

  # outcome: employment 2014 (cached panel) and 2017 (Wayback-fetched xlsx, bespoke header layout)
  panel <- readRDS("raw_data/01_bls_panel.rds")
  emp_2014 <- panel %>% filter(year == 2014) %>% transmute(fips, employed_2014 = employed)

  wayback_url <- "https://web.archive.org/web/20190111093926/https://www.bls.gov/lau/laucnty17.xlsx"
  if (!file.exists("raw_data/laucnty17.xlsx")) {
    GET(wayback_url, write_disk("raw_data/laucnty17.xlsx", overwrite = TRUE), timeout(60))
  }
  emp_2017_raw <- read_excel("raw_data/laucnty17.xlsx", col_names = FALSE, skip = 6)
  names(emp_2017_raw)[1:10] <- c("laus_code", "state_fips", "county_fips", "county_name", "year",
                                  "blank", "labor_force", "employed", "unemployed", "unemployment_rate")
  emp_2017 <- emp_2017_raw %>%
    filter(!is.na(state_fips), !is.na(county_fips)) %>%
    transmute(fips = paste0(sprintf("%02s", state_fips), sprintf("%03s", county_fips)),
              employed_2017 = as.numeric(employed))
  outcome_placebo <- emp_2014 %>% inner_join(emp_2017, by = "fips") %>%
    mutate(log_outcome = log(employed_2017 / employed_2014))

  # proprietor_share_2013 (cached BEA CAEMP25N bulk file, already has a 2013 column)
  caemp <- read_csv("raw_data/CAEMP25N_extracted/CAEMP25N__ALL_AREAS_2001_2022.csv", col_types = cols(.default = "c"))
  proprietor_2013 <- caemp %>%
    mutate(fips = trimws(GeoFIPS)) %>%
    filter(nchar(fips) == 5, substr(fips, 3, 5) != "000", LineCode %in% c("10", "40")) %>%
    transmute(fips, LineCode, emp = as.numeric(`2013`)) %>%
    pivot_wider(names_from = LineCode, values_from = emp, names_prefix = "line_") %>%
    mutate(proprietor_share_2013 = line_40 / line_10) %>%
    select(fips, proprietor_share_2013)

  # migration_rate_2013 (fresh IRS SOI 2012-2013 pull, same "code 96" logic as pipeline/12)
  download_soi <- function(fname) {
    path <- paste0("raw_data/", fname)
    if (!file.exists(path)) GET(paste0("https://www.irs.gov/pub/irs-soi/", fname),
                                 add_headers(`User-Agent` = ua), write_disk(path, overwrite = TRUE), timeout(60))
    path
  }
  read_soi_csv <- function(path) read_csv(path, col_types = cols(.default = "c"))
  inflow_1213 <- read_soi_csv(download_soi("countyinflow1213.csv")) %>%
    filter(y1_statefips == "96", y1_countyfips == "000") %>%
    transmute(fips = paste0(y2_statefips, y2_countyfips), inflow_n2 = as.numeric(n2))
  outflow_1213 <- read_soi_csv(download_soi("countyoutflow1213.csv")) %>%
    filter(y2_statefips == "96", y2_countyfips == "000") %>%
    transmute(fips = paste0(y1_statefips, y1_countyfips), outflow_n2 = as.numeric(n2))

  # population/density/pop_trend 2013 (cached PEP 2010s file + cached Gazetteer)
  pep <- read_csv("raw_data/co-est2019-alldata.csv", show_col_types = FALSE,
                   col_types = cols(.default = "c"), locale = locale(encoding = "latin1"))
  pop_2013 <- pep %>% filter(SUMLEV == "050") %>%
    transmute(fips = paste0(sprintf("%02s", STATE), sprintf("%03s", COUNTY)),
              pop_2010 = as.numeric(POPESTIMATE2010), pop_2013 = as.numeric(POPESTIMATE2013))
  gazetteer <- read_tsv(list.files("raw_data/gaz2024", pattern = "\\.txt$", full.names = TRUE)[1],
                         show_col_types = FALSE, col_types = cols(GEOID = col_character())) %>%
    transmute(fips = sprintf("%05s", GEOID), land_area_sqmi = ALAND_SQMI)
  pop_density_2013 <- pop_2013 %>% left_join(gazetteer, by = "fips") %>%
    transmute(fips, population_2013 = pop_2013, population_density_2013 = pop_2013 / land_area_sqmi,
              pop_trend_2013 = 100 * ((pop_2013 / pop_2010)^(1/3) - 1))
  migration_2013 <- inflow_1213 %>% inner_join(outflow_1213, by = "fips") %>%
    inner_join(pop_2013 %>% select(fips, pop_2013), by = "fips") %>%
    mutate(net_migration_rate_2013 = 1000 * (inflow_n2 - outflow_n2) / pop_2013) %>%
    select(fips, net_migration_rate_2013)

  # unemployment_rate_2013 (clean-format xlsx, same header-detection as pipeline/18)
  preview <- read_excel("raw_data/laucnty13.xlsx", col_names = FALSE, n_max = 10)
  header_row <- which(apply(preview, 1, function(r) any(grepl("Unemployment Rate", r, ignore.case = TRUE))))
  raw13 <- read_excel("raw_data/laucnty13.xlsx", skip = header_row - 1)
  names(raw13) <- make.unique(names(raw13))
  unemployment_2013 <- raw13 %>%
    rename_with(~"state_fips", matches("State FIPS", ignore.case = TRUE)) %>%
    rename_with(~"county_fips", matches("County FIPS", ignore.case = TRUE)) %>%
    rename_with(~"rate_raw", matches("Unemployment Rate", ignore.case = TRUE)) %>%
    filter(!is.na(state_fips), !is.na(county_fips)) %>%
    transmute(fips = paste0(sprintf("%02s", state_fips), sprintf("%03s", county_fips)),
              pre_unemployment_2013 = as.numeric(rate_raw))

  # working_age_share_2013 (cached per-state agesex files, YEAR code 6 = 2013)
  pull_agesex13 <- function(sf) {
    read_csv(sprintf("raw_data/cc-est2019-agesex-%s.csv", sf), show_col_types = FALSE, col_types = cols(.default = "c")) %>%
      filter(SUMLEV == "050", YEAR == "6") %>%
      transmute(fips = paste0(sprintf("%02s", STATE), sprintf("%03s", COUNTY)),
                pop_est = as.numeric(POPESTIMATE),
                working_age_pop = as.numeric(AGE2544_TOT) + as.numeric(AGE4564_TOT))
  }
  agesex_2013 <- bind_rows(lapply(valid_state_fips, pull_agesex13)) %>%
    mutate(working_age_share_2013 = 100 * working_age_pop / pop_est) %>%
    select(fips, working_age_share_2013)

  # rucc_2013 (cached xls, already pulled for the COVID side)
  rucc_2013 <- read_excel("raw_data/rucc2013.xls") %>%
    transmute(fips = sprintf("%05s", FIPS), rucc_2013 = as.integer(RUCC_2013))

  # SAIPE 2013 (income + poverty) -- fresh pull, only new Census API call needed
  url <- paste0("https://api.census.gov/data/timeseries/poverty/saipe?get=NAME,SAEPOVRTALL_PT,SAEMHI_PT&for=county:*&in=state:*&YEAR=2013&key=", census_key)
  raw_saipe <- fromJSON(url)
  saipe_2013 <- as_tibble(raw_saipe[-1, ], .name_repair = "minimal")
  names(saipe_2013) <- raw_saipe[1, ]
  saipe_2013 <- saipe_2013 %>%
    transmute(fips = paste0(state, county), poverty_rate_2013 = as.numeric(SAEPOVRTALL_PT),
              median_hh_income_2013 = as.numeric(SAEMHI_PT))

  pieces <- list(outcome_placebo = outcome_placebo, proprietor = proprietor_2013, migration = migration_2013,
                 pop_density = pop_density_2013, unemployment = unemployment_2013, agesex = agesex_2013,
                 rucc = rucc_2013, saipe = saipe_2013)
  saveRDS(pieces, placebo_cache)
}
```

```{r placebo-model}
placebo_df <- pieces$outcome_placebo %>%
  left_join(pieces$proprietor, by = "fips") %>%
  left_join(pieces$migration, by = "fips") %>%
  left_join(pieces$unemployment, by = "fips") %>%
  left_join(pieces$agesex, by = "fips") %>%
  left_join(pieces$saipe, by = "fips") %>%
  left_join(pieces$pop_density, by = "fips") %>%
  left_join(pieces$rucc, by = "fips") %>%
  filter(if_all(c(proprietor_share_2013, net_migration_rate_2013, pre_unemployment_2013, working_age_share_2013,
                   median_hh_income_2013, poverty_rate_2013, population_2013, population_density_2013,
                   rucc_2013, pop_trend_2013), ~ !is.na(.))) %>%
  mutate(
    log_population = log(population_2013), log_density = log(population_density_2013),
    rucc = relevel(factor(rucc_2013), ref = "1"),
    proprietor_share_z = as.numeric(scale(proprietor_share_2013)),
    migration_rate_z = as.numeric(scale(net_migration_rate_2013)),
    unemployment_rate_z = as.numeric(scale(pre_unemployment_2013)),
    working_age_share_z = as.numeric(scale(working_age_share_2013)),
    income = median_hh_income_2013, poverty_rate = poverty_rate_2013, pop_trend = pop_trend_2013
  )

cat("Placebo complete-case n:", nrow(placebo_df), "\n")

placebo_formula <- as.formula(paste("log_outcome ~", paste(c(studied_z, controls), collapse = " + ")))
placebo_fits_cache <- "raw_data/modeling_andrew_placebo_fits.rds"
if (file.exists(placebo_fits_cache)) {
  placebo_fits <- readRDS(placebo_fits_cache)
  placebo_ols <- placebo_fits$placebo_ols
  placebo_rf <- placebo_fits$placebo_rf
} else {
  placebo_ols <- lm_robust(placebo_formula, data = placebo_df, se_type = "HC1")
  set.seed(20260729)
  placebo_rf <- randomForest(placebo_formula, data = placebo_df, ntree = 500, importance = FALSE)
  saveRDS(list(placebo_ols = placebo_ols, placebo_rf = placebo_rf), placebo_fits_cache)
}

as_tibble(summary(placebo_ols)$coefficients[studied_z, ], rownames = "term") %>%
  rename(estimate = Estimate, robust_se = `Std. Error`, p = `Pr(>|t|)`) %>%
  select(term, estimate, robust_se, p)
```

```{r placebo-fit-stats}
tibble(
  model = c("2008 (primary)", "COVID (primary)", "Placebo (2014-2017)"),
  oob_r_squared = c(res_2008$rf_a$rsq[500], res_covid$rf_a$rsq[500], placebo_rf$rsq[500]),
  oob_rmse = c(sqrt(res_2008$rf_a$mse[500]), sqrt(res_covid$rf_a$mse[500]), sqrt(placebo_rf$mse[500])),
  ols_adj_r_squared = c(res_2008$ols$adj.r.squared, res_covid$ols$adj.r.squared, placebo_ols$adj.r.squared)
)
```

## Relaxed LASSO

Same spec as the standard LASSO section (same 4 studied predictors, same 6
controls incl. all 8 RUCC dummies unpenalized, same 10-fold CV, same seed) --
`relax = TRUE` fits, at every lambda, both the usual shrunk lasso fit and an
unpenalized refit on that lambda's active set, then blends the two by
`gamma` (gamma=1 is the standard shrunk lasso, gamma=0 is the fully
de-biased refit). CV picks `lambda` and `gamma` jointly.

```{r relax-lasso-cache}
relax_cache <- "raw_data/modeling_andrew_relaxed_lasso.rds"

run_relaxed_lasso <- function(res) {
  df <- res$df
  X <- model.matrix(as.formula(paste("~", paste(c(studied_z, controls), collapse = " + "))), data = df)[, -1]
  y <- df$log_outcome
  penalty <- ifelse(colnames(X) %in% studied_z, 1, 0)
  set.seed(20260729)
  cv.glmnet(X, y, alpha = 1, nfolds = 10, standardize = TRUE, penalty.factor = penalty, relax = TRUE)
}

if (file.exists(relax_cache)) {
  relaxed <- readRDS(relax_cache)
} else {
  relaxed <- list("2008" = run_relaxed_lasso(res_2008), "covid" = run_relaxed_lasso(res_covid))
  saveRDS(relaxed, relax_cache)
}
relax_2008 <- relaxed[["2008"]]
relax_covid <- relaxed[["covid"]]
```

### Best gamma and lambda from CV

```{r relax-best-tuning}
tibble(
  crisis = c("2008", "covid"),
  lambda_min = c(relax_2008$lambda.min, relax_covid$lambda.min),
  gamma_min = c(relax_2008$relaxed$gamma.min, relax_covid$relaxed$gamma.min),
  lambda_1se = c(relax_2008$lambda.1se, relax_covid$lambda.1se),
  gamma_1se = c(relax_2008$relaxed$gamma.1se, relax_covid$relaxed$gamma.1se)
)
```

### Coefficients at lambda.min and lambda.1se

```{r relax-coefs}
relax_coef <- function(res, s, gamma) as.matrix(coef(res, s = s, gamma = gamma))[, 1]

relax_2008_min <- relax_coef(relax_2008, relax_2008$lambda.min, relax_2008$relaxed$gamma.min)
relax_2008_1se <- relax_coef(relax_2008, relax_2008$lambda.1se, relax_2008$relaxed$gamma.1se)
relax_covid_min <- relax_coef(relax_covid, relax_covid$lambda.min, relax_covid$relaxed$gamma.min)
relax_covid_1se <- relax_coef(relax_covid, relax_covid$lambda.1se, relax_covid$relaxed$gamma.1se)

tibble(term = names(relax_2008_min), coef_2008_min = relax_2008_min, coef_2008_1se = relax_2008_1se) %>%
  left_join(tibble(term = names(relax_covid_min), coef_covid_min = relax_covid_min, coef_covid_1se = relax_covid_1se), by = "term")
```

### Standard LASSO vs. relaxed LASSO vs. OLS -- studied predictors only

```{r relax-comparison-table}
ols_coef <- function(res) summary(res$ols)$coefficients[studied_z, "Estimate"]

three_way <- bind_rows(
  tibble(term = studied_z, crisis = "2008", lambda_level = "min",
         standard_lasso = res_2008$coef_min[studied_z], relaxed_lasso = relax_2008_min[studied_z], ols = ols_coef(res_2008)),
  tibble(term = studied_z, crisis = "2008", lambda_level = "1se",
         standard_lasso = res_2008$coef_1se[studied_z], relaxed_lasso = relax_2008_1se[studied_z], ols = ols_coef(res_2008)),
  tibble(term = studied_z, crisis = "covid", lambda_level = "min",
         standard_lasso = res_covid$coef_min[studied_z], relaxed_lasso = relax_covid_min[studied_z], ols = ols_coef(res_covid)),
  tibble(term = studied_z, crisis = "covid", lambda_level = "1se",
         standard_lasso = res_covid$coef_1se[studied_z], relaxed_lasso = relax_covid_1se[studied_z], ols = ols_coef(res_covid))
) %>%
  mutate(
    dist_standard_to_ols = abs(standard_lasso - ols),
    dist_relaxed_to_ols = abs(relaxed_lasso - ols),
    relaxed_closer_to_ols = dist_relaxed_to_ols < dist_standard_to_ols,
    newly_nonzero_under_relax = standard_lasso == 0 & relaxed_lasso != 0
  )
three_way %>% select(term, crisis, lambda_level, standard_lasso, relaxed_lasso, ols, relaxed_closer_to_ols, newly_nonzero_under_relax)
```

### CV error curves

```{r relax-cv-plot-2008, fig.width=7, fig.height=5}
plot(relax_2008)
title("2008: relaxed LASSO CV error by lambda and gamma", line = 3)
```

```{r relax-cv-plot-covid, fig.width=7, fig.height=5}
plot(relax_covid)
title("COVID: relaxed LASSO CV error by lambda and gamma", line = 3)
```

## Presentation figures

All seven figures below are built **only from objects already saved to
disk** by the sections above (`modeling_andrew_results.rds`,
`modeling_andrew_relaxed_lasso.rds`, `modeling_andrew_wealthband_fits.rds`,
`modeling_andrew_placebo_fits.rds`) -- no model is refit here. (Two of those
caches -- wealth-band and placebo fits -- didn't exist as standalone files
before this section; they were only ever computed inline in Sections A/B
above. They've been added to those sections' existing `file.exists()` cache
pattern, consistent with every other model in this document, so this section
itself only ever reads, never fits.)

Colorblind-safe palette (validated categorical set): blue `#2a78d6`, orange
`#eb6834`, aqua `#1baf7a` -- the first three slots, which validate against
every pairing including when several appear in the same panel at once.
A true 4th categorical hue (yellow) is avoided by design: paired with orange
in an all-series-visible chart it fails the accessibility floor, so the
`lasso_path` plot's 4th line uses ink + a dashed linetype instead of a 4th
hue. Wealth bands use a single-hue **sequential** ramp (light->dark blue)
rather than 3 unrelated categorical hues, since the bands are ordinal
(lower < middle < higher), not nominal.

```{r figures-setup}
library(ggplot2)

dir.create("figures", showWarnings = FALSE)

studied_labels <- c(proprietor_share_z = "Self-employment", migration_rate_z = "Migration rate",
                     unemployment_rate_z = "Unemployment rate", working_age_share_z = "Working-age share")
pal <- c(blue = "#2a78d6", orange = "#eb6834", aqua = "#1baf7a", yellow = "#eda100")
seq_blue <- c(lower = "#86b6ef", middle = "#2a78d6", higher = "#104281")

# base_size 14+, minimal gridlines (vertical only, hairline), no in-plot
# title (titles live on the slide, not the PNG), legend below
theme_pub <- function(base_size = 14) {
  theme_minimal(base_size = base_size) +
    theme(panel.grid.minor = element_blank(), panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_line(color = "#e1e0d9", linewidth = 0.3),
          strip.text = element_text(face = "bold", size = rel(0.95)), plot.title = element_blank(),
          legend.position = "bottom", axis.title = element_text(size = rel(0.9)))
}

ols_ci_table <- function(res, crisis_label) {
  s <- summary(res$ols)$coefficients[studied_z, ]
  tibble(term = studied_z, estimate = s[, "Estimate"], ci_low = s[, "CI Lower"], ci_high = s[, "CI Upper"], crisis = crisis_label)
}
```

### 1. coefficient_plot.png

```{r fig-coefficient-plot, fig.width=10, fig.height=6}
coef_data <- bind_rows(ols_ci_table(res_2008, "2008"), ols_ci_table(res_covid, "COVID")) %>%
  mutate(predictor = factor(studied_labels[term], levels = rev(studied_labels)))

p1 <- ggplot(coef_data, aes(x = estimate, y = predictor)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#898781", linewidth = 0.5) +
  geom_pointrange(aes(xmin = ci_low, xmax = ci_high), color = unname(pal["blue"]), linewidth = 0.8, size = 0.6) +
  facet_wrap(~crisis) +
  labs(x = "OLS coefficient (log recovery ratio per 1 SD), 95% CI (HC1 robust SE)", y = NULL) +
  theme_pub()
ggsave("figures/coefficient_plot.png", p1, width = 10, height = 6, dpi = 300)
p1
```

### 2. method_agreement.png

```{r fig-method-agreement, fig.width=10, fig.height=6}
relax_coef_vec <- function(res, s, gamma) as.matrix(coef(res, s = s, gamma = gamma))[, 1]
relax_2008_min <- relax_coef_vec(relax_2008, relax_2008$lambda.min, relax_2008$relaxed$gamma.min)
relax_covid_min <- relax_coef_vec(relax_covid, relax_covid$lambda.min, relax_covid$relaxed$gamma.min)

method_data <- bind_rows(
  ols_ci_table(res_2008, "2008") %>% mutate(method = "OLS"),
  ols_ci_table(res_covid, "COVID") %>% mutate(method = "OLS"),
  tibble(term = studied_z, estimate = res_2008$coef_min[studied_z], ci_low = NA, ci_high = NA, crisis = "2008", method = "Standard LASSO"),
  tibble(term = studied_z, estimate = res_covid$coef_min[studied_z], ci_low = NA, ci_high = NA, crisis = "COVID", method = "Standard LASSO"),
  tibble(term = studied_z, estimate = relax_2008_min[studied_z], ci_low = NA, ci_high = NA, crisis = "2008", method = "Relaxed LASSO"),
  tibble(term = studied_z, estimate = relax_covid_min[studied_z], ci_low = NA, ci_high = NA, crisis = "COVID", method = "Relaxed LASSO")
) %>%
  mutate(predictor = factor(studied_labels[term], levels = rev(studied_labels)),
         method = factor(method, levels = c("OLS", "Standard LASSO", "Relaxed LASSO")))

dodge_width <- 0.6
method_colors <- c("OLS" = unname(pal["blue"]), "Standard LASSO" = unname(pal["orange"]), "Relaxed LASSO" = unname(pal["aqua"]))
p2 <- ggplot(method_data, aes(x = estimate, y = predictor, color = method, shape = method)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#898781", linewidth = 0.5) +
  geom_errorbar(aes(xmin = ci_low, xmax = ci_high), width = 0.15, linewidth = 0.6, orientation = "y",
                position = position_dodge(width = dodge_width), na.rm = TRUE) +
  geom_point(size = 3, position = position_dodge(width = dodge_width)) +
  scale_color_manual(values = method_colors, name = NULL) +
  scale_shape_manual(values = c("OLS" = 16, "Standard LASSO" = 17, "Relaxed LASSO" = 15), name = NULL) +
  facet_wrap(~crisis) +
  labs(x = "Coefficient (lambda.min for LASSO variants)", y = NULL) +
  theme_pub()
ggsave("figures/method_agreement.png", p2, width = 10, height = 6, dpi = 300)
p2
```

### 3. lasso_path.png

Excludes controls (unpenalized, flat lines add no information). Only 3
predictors get a categorical hue -- see the palette note above for why the
4th (working-age share) is ink + dashed instead of a 4th hue.

```{r fig-lasso-path, fig.width=10, fig.height=6}
path_line_colors <- c("Self-employment" = unname(pal["blue"]), "Migration rate" = unname(pal["orange"]),
                       "Unemployment rate" = unname(pal["aqua"]), "Working-age share" = "#2b2b2b")
path_linetypes <- c("Self-employment" = "solid", "Migration rate" = "solid",
                     "Unemployment rate" = "solid", "Working-age share" = "dashed")

extract_path <- function(res, crisis_label) {
  co <- as.matrix(coef(res$cvfit$glmnet.fit))
  lambda <- res$cvfit$glmnet.fit$lambda
  map_dfr(studied_z, function(v) {
    tibble(term = v, predictor = studied_labels[v], log_lambda = log(lambda), coef = co[v, ], crisis = crisis_label)
  }) %>%
    mutate(lambda_min = res$cvfit$lambda.min, lambda_1se = res$cvfit$lambda.1se)
}
path_data <- bind_rows(extract_path(res_2008, "2008"), extract_path(res_covid, "COVID"))

# direct labels at the left edge, positioned using EACH FACET'S OWN data
# range (facets use free_x scales with very different lambda ranges)
label_data <- path_data %>% group_by(crisis, predictor) %>% slice_min(log_lambda, n = 1) %>% ungroup()
facet_ranges <- path_data %>% group_by(crisis) %>% summarise(x_range = max(log_lambda) - min(log_lambda), .groups = "drop")
label_data <- label_data %>% left_join(facet_ranges, by = "crisis") %>% mutate(label_x = log_lambda - 0.03 * x_range)

vline_data <- path_data %>% distinct(crisis, lambda_min, lambda_1se) %>%
  pivot_longer(c(lambda_min, lambda_1se), names_to = "which", values_to = "lambda") %>%
  mutate(log_lambda = log(lambda))

p3 <- ggplot(path_data, aes(x = log_lambda, y = coef, color = predictor, linetype = predictor)) +
  geom_vline(data = vline_data, aes(xintercept = log_lambda), color = "#898781", linewidth = 0.4, linetype = "dotted") +
  geom_hline(yintercept = 0, color = "#c3c2b7", linewidth = 0.4) +
  geom_line(linewidth = 0.9) +
  geom_text(data = label_data, aes(x = label_x, y = coef, label = predictor), hjust = 1, size = 3.6, show.legend = FALSE) +
  scale_color_manual(values = path_line_colors, guide = "none") +
  scale_linetype_manual(values = path_linetypes, guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0.42, 0.05))) +
  facet_wrap(~crisis, scales = "free_x") +
  labs(x = "log(lambda) -- dotted lines mark lambda.min and lambda.1se", y = "Standardized coefficient") +
  theme_pub()
ggsave("figures/lasso_path.png", p3, width = 10, height = 6, dpi = 300)
p3
```

### 4. rf_importance.png

Model 3b only (no income/poverty) -- model 3a's importance is deliberately
never computed, per the rule in that section.

```{r fig-rf-importance, fig.width=10, fig.height=6}
rf_imp_table <- function(res, crisis_label) {
  imp <- importance(res$rf_b, type = 1)
  var_labels <- c(studied_labels, log_population = "log(population)", log_density = "log(density)",
                   rucc = "RUCC", pop_trend = "Population trend")
  tibble(term = rownames(imp), pct_inc_mse = imp[, 1], crisis = crisis_label) %>%
    mutate(label = var_labels[term], is_studied = term %in% studied_z)
}
rf_data <- bind_rows(rf_imp_table(res_2008, "2008"), rf_imp_table(res_covid, "COVID"))
# fct_reorder() alone shares ONE bar order across facets -- key on
# label+crisis so each panel sorts by its OWN values, then strip the
# crisis suffix back off the displayed tick labels
rf_data <- rf_data %>% mutate(label_key = fct_reorder(paste0(label, "___", crisis), pct_inc_mse))

p4 <- ggplot(rf_data, aes(x = pct_inc_mse, y = label_key, fill = is_studied)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = c("TRUE" = unname(pal["blue"]), "FALSE" = "#c3c2b7"), guide = "none") +
  scale_y_discrete(labels = function(x) sub("___.*", "", x)) +
  facet_wrap(~crisis, scales = "free_y") +
  labs(x = "Permutation importance (% increase in MSE when shuffled)", y = NULL) +
  theme_pub()
ggsave("figures/rf_importance.png", p4, width = 10, height = 6, dpi = 300)
p4
```

### 5. predictability.png

```{r fig-predictability, fig.width=10, fig.height=6}
pred_data <- tibble(
  model = factor(c("2008", "COVID", "Placebo\n(2014-2017)"), levels = c("2008", "COVID", "Placebo\n(2014-2017)")),
  oob_r_squared = c(res_2008$rf_a$rsq[500], res_covid$rf_a$rsq[500], placebo_rf$rsq[500]),
  oob_rmse = c(sqrt(res_2008$rf_a$mse[500]), sqrt(res_covid$rf_a$mse[500]), sqrt(placebo_rf$mse[500]))
) %>% pivot_longer(-model, names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(metric, oob_r_squared = "OOB R-squared", oob_rmse = "OOB RMSE (log scale)"))

model_colors <- c("2008" = unname(pal["blue"]), "COVID" = unname(pal["orange"]), "Placebo\n(2014-2017)" = unname(pal["aqua"]))
p5 <- ggplot(pred_data, aes(x = model, y = value, fill = model)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = ifelse(metric == "OOB R-squared", scales::percent(value, accuracy = 0.1), sprintf("%.3f", value))),
            vjust = -0.5, size = 4) +
  scale_fill_manual(values = model_colors, guide = "none") +
  facet_wrap(~metric, scales = "free_y") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = NULL, y = NULL) +
  theme_pub()
ggsave("figures/predictability.png", p5, width = 10, height = 6, dpi = 300)
p5
```

### 6. outcome_distributions.png

```{r fig-outcome-distributions, fig.width=10, fig.height=6}
outcome_data <- bind_rows(
  tibble(log_ratio = log(res_2008$df$outcome_raw), crisis = "2008"),
  tibble(log_ratio = log(res_covid$df$outcome_raw), crisis = "COVID")
)
sd_labels <- outcome_data %>% group_by(crisis) %>%
  summarise(sd_val = sd(log_ratio), x_pos = max(log_ratio) * 0.6, .groups = "drop")

p6 <- ggplot(outcome_data, aes(x = log_ratio)) +
  geom_histogram(aes(y = after_stat(density)), bins = 60, fill = unname(pal["blue"]), alpha = 0.85) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#2b2b2b", linewidth = 0.6) +
  geom_text(data = sd_labels, aes(x = x_pos, y = Inf, label = sprintf("SD = %.3f", sd_val)),
            vjust = 1.5, size = 4.2, inherit.aes = FALSE) +
  facet_wrap(~crisis) +
  labs(x = "log(recovery ratio) -- 0 marks exactly recovered", y = "Density") +
  theme_pub()
ggsave("figures/outcome_distributions.png", p6, width = 10, height = 6, dpi = 300)
p6
```

### 7. wealth_bands.png

Band coefficients use a single-hue **sequential** ramp (light->dark),
since the three bands are ordinal, not an unordered category set.

```{r fig-wealth-bands, fig.width=10, fig.height=6}
bands <- c("lower", "middle", "higher")
band_table <- imap_dfr(band_fits, function(crisis_fits, crisis) {
  imap_dfr(crisis_fits, function(fit, band) {
    s <- summary(fit)$coefficients[studied_z, ]
    as_tibble(s, rownames = "term") %>%
      rename(estimate = Estimate, ci_low = `CI Lower`, ci_high = `CI Upper`) %>%
      mutate(crisis = recode(crisis, "2008" = "2008", "covid" = "COVID"), band = band)
  })
}) %>%
  mutate(predictor = factor(studied_labels[term], levels = rev(studied_labels)),
         band = factor(band, levels = bands, labels = c("Lower third", "Middle third", "Higher third")))

dodge_w <- 0.6
p7 <- ggplot(band_table, aes(x = estimate, y = predictor, color = band)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#898781", linewidth = 0.5) +
  geom_errorbar(aes(xmin = ci_low, xmax = ci_high), width = 0.15, linewidth = 0.7,
                orientation = "y", position = position_dodge(width = dodge_w)) +
  geom_point(size = 3, position = position_dodge(width = dodge_w)) +
  scale_color_manual(values = c("Lower third" = unname(seq_blue["lower"]), "Middle third" = unname(seq_blue["middle"]),
                                 "Higher third" = unname(seq_blue["higher"])), name = "Income tercile") +
  facet_wrap(~crisis) +
  labs(x = "OLS coefficient (log recovery ratio per 1 SD), 95% CI (HC1 robust SE)", y = NULL) +
  theme_pub()
ggsave("figures/wealth_bands.png", p7, width = 10, height = 6, dpi = 300)
p7
```

### 8. relaxed_lasso.png

Relaxed LASSO alone (no OLS/standard-LASSO overlay -- that comparison is
already `method_agreement.png`). Reuses `relax_2008_min` / `relax_covid_min`,
already computed in the method-agreement chunk above from the cached
`modeling_andrew_relaxed_lasso.rds` -- no refit here.

```{r fig-relaxed-lasso, fig.width=10, fig.height=6}
relax_solo_data <- bind_rows(
  tibble(term = studied_z, estimate = relax_2008_min[studied_z], crisis = "2008"),
  tibble(term = studied_z, estimate = relax_covid_min[studied_z], crisis = "COVID")
) %>%
  mutate(predictor = factor(studied_labels[term], levels = rev(studied_labels)))

p8 <- ggplot(relax_solo_data, aes(x = estimate, y = predictor)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#898781", linewidth = 0.5) +
  geom_point(color = unname(pal["blue"]), size = 3.2) +
  geom_text(aes(label = sprintf("%.4f", estimate)), vjust = -1.3, size = 3.8) +
  facet_wrap(~crisis) +
  scale_x_continuous(expand = expansion(mult = c(0.25, 0.25))) +
  labs(x = "Relaxed LASSO coefficient (lambda.min)", y = NULL) +
  theme_pub()
ggsave("figures/relaxed_lasso.png", p8, width = 10, height = 6, dpi = 300)
p8
```

### 9. lasso_cv_curve.png

Standard (non-relaxed) LASSO's CV curve -- the classic `plot(cv.glmnet(...))`
shape, rebuilt in ggplot to match this deck's style. Reads `res_2008$cvfit` /
`res_covid$cvfit`, already fit in the `run-crisis-models` chunk from cached
`modeling_andrew_results.rds` -- no refit here. Only the 4 studied predictors
are penalized (`penalty.factor = 0` on all 6 controls), so `nzero` only
ranges 13-17 across the whole path, not down to 0 -- the curve is flat until
the last studied predictors get dropped, unlike a fully-penalized textbook
example.

```{r fig-lasso-cv-curve, fig.width=10, fig.height=6}
extract_cv_curve <- function(res, crisis_label) {
  cv <- res$cvfit
  tibble(log_lambda = log(cv$lambda), cvm = cv$cvm, cvup = cv$cvup, cvlo = cv$cvlo,
         nzero = cv$nzero, crisis = crisis_label) %>%
    mutate(lambda_min = cv$lambda.min, lambda_1se = cv$lambda.1se)
}
cv_curve_data <- bind_rows(extract_cv_curve(res_2008, "2008"), extract_cv_curve(res_covid, "COVID"))

cv_vline_data <- cv_curve_data %>% distinct(crisis, lambda_min, lambda_1se) %>%
  pivot_longer(c(lambda_min, lambda_1se), names_to = "which", values_to = "lambda") %>%
  mutate(log_lambda = log(lambda))

# nonzero-predictor count labeled every 5th lambda along the top, echoing the
# base-R plot.cv.glmnet() convention (top-axis nzero counts)
nzero_labels <- cv_curve_data %>% group_by(crisis) %>% slice(seq(1, n(), by = 5)) %>% ungroup()

p9 <- ggplot(cv_curve_data, aes(x = log_lambda, y = cvm)) +
  geom_vline(data = cv_vline_data, aes(xintercept = log_lambda), color = "#898781", linewidth = 0.4, linetype = "dotted") +
  geom_errorbar(aes(ymin = cvlo, ymax = cvup), color = "#c3c2b7", width = 0, linewidth = 0.4) +
  geom_point(color = unname(pal["blue"]), size = 1.6) +
  geom_text(data = nzero_labels, aes(label = nzero), y = Inf, vjust = 1.6, size = 3, color = "#5b5a54") +
  facet_wrap(~crisis, scales = "free") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.18))) +
  labs(x = "log(lambda) -- dotted lines mark lambda.min and lambda.1se; numbers = predictors kept",
       y = "Mean-Squared Error (10-fold CV)") +
  theme_pub()
ggsave("figures/lasso_cv_curve.png", p9, width = 10, height = 6, dpi = 300)
p9
```
