---
title: "Option C: Expanded Predictors via ACS (Race, Education, Industry Mix)"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    df-print: kable
execute:
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(estimatr)   # lm_robust, HC1 SEs
library(glmnet)
library(randomForest)
library(jsonlite)
library(httr)

set.seed(20260729)

dir.create("raw_data", showWarnings = FALSE)
dir.create("figures_optionc", showWarnings = FALSE)
census_key <- Sys.getenv("CENSUS_API_KEY")
```

This is a **separate extension** of Option A, not a replacement. Option A
(`modeling_andrew.qmd`, its cached fits, and `figures/*.png`) is untouched --
read-only everywhere it's referenced below. This document builds its own
`spine_data_optionc.csv` and writes any figures to `figures_optionc/`.

Option A kept: 4 studied predictors (self-employment, migration, unemployment,
working-age share), same 6 controls, same log outcome, same RUCC-as-factor
spec, same seed. This document adds three predictor buckets on top -- race,
education, industry mix -- built from ACS/PEP sources not already in the
pipeline, then re-runs the full Option A model set on the resulting (smaller)
sample.

## 1. New predictor: RACE (Census PEP, not ACS)

Race is a full-coverage PEP series, not ACS -- every county gets a value,
unlike education/industry below. Built as shares of total population, using
the Census "race alone" categories (White/Black/AIAN/Asian/NHPI/Two-or-more),
stable OMB categories across both vintages, plus a separate Hispanic-origin
share (an ethnicity flag orthogonal to race in Census methodology, not a
7th race category).

- **2007**: `co-est00int-sexracehisp.csv` -- a single national county-level
  file from the 2000-2010 intercensal series, already broken out by
  SEX x ORIGIN x RACE with a direct `POPESTIMATE2007` column. Simpler than
  the per-state single-year-age `co-est00int-alldata-<state>.csv` files (not
  needed here since no age breakdown is wanted) -- confirmed by request,
  `SEX=0` (both sexes), `RACE=0`/`ORIGIN=0` total = 52,405 for Autauga
  County, AL, matching the population already used elsewhere in this
  project's pipeline for the same county/year.
- **2019**: `cc-est2019-alldata.csv` -- the single national CC-EST2019 file
  with a full race/Hispanic-origin breakdown by single-year age (176MB;
  the per-state files already cached elsewhere in this project's
  `raw_data/` are the age-sex-only variant and don't carry race). Filtered
  to `AGEGRP == "0"` (all ages) and `YEAR == "12"`, confirmed to reproduce
  the already-known 2019 population for Autauga County (55,869) exactly.

```{r race-2007, cache=TRUE}
race_2007_path <- "raw_data/optionc_race_2007.rds"
if (file.exists(race_2007_path)) {
  race_2007 <- readRDS(race_2007_path)
} else {
  sexracehisp_path <- "raw_data/co-est00int-sexracehisp.csv"
  if (!file.exists(sexracehisp_path)) {
    GET("https://www2.census.gov/programs-surveys/popest/datasets/2000-2010/intercensal/county/co-est00int-sexracehisp.csv",
        add_headers(`User-Agent` = "Mozilla/5.0"), write_disk(sexracehisp_path, overwrite = TRUE), timeout(120))
  }
  d07 <- read_csv(sexracehisp_path, show_col_types = FALSE, col_types = cols(.default = "c")) %>%
    filter(SUMLEV == "050", SEX == "0") %>%
    transmute(fips = paste0(sprintf("%02s", STATE), sprintf("%03s", COUNTY)),
              ORIGIN, RACE, pop2007 = as.numeric(POPESTIMATE2007))

  race07 <- d07 %>% filter(ORIGIN == "0") %>% select(fips, RACE, pop2007) %>%
    pivot_wider(names_from = RACE, values_from = pop2007, names_prefix = "race_")
  hisp07 <- d07 %>% filter(RACE == "0") %>% select(fips, ORIGIN, pop2007) %>%
    pivot_wider(names_from = ORIGIN, values_from = pop2007, names_prefix = "origin_")

  # RACE codes: 0=total, 1=White, 2=Black, 3=AIAN, 4=Asian, 5=NHPI, 6=Two-or-more (all "alone", any origin)
  race_2007 <- race07 %>% left_join(hisp07, by = "fips") %>%
    transmute(fips, total_2007 = race_0,
      white_share_2007 = 100 * race_1 / race_0, black_share_2007 = 100 * race_2 / race_0,
      aian_share_2007 = 100 * race_3 / race_0, asian_share_2007 = 100 * race_4 / race_0,
      nhpi_share_2007 = 100 * race_5 / race_0, tom_share_2007 = 100 * race_6 / race_0,
      hispanic_share_2007 = 100 * origin_2 / race_0)
  saveRDS(race_2007, race_2007_path)
}
cat("2007 race table:", nrow(race_2007), "counties (full PEP coverage, no ACS threshold)\n")
```

```{r race-2019, cache=TRUE}
race_2019_path <- "raw_data/optionc_race_2019.rds"
if (file.exists(race_2019_path)) {
  race_2019 <- readRDS(race_2019_path)
} else {
  alldata_path <- "raw_data/cc-est2019-alldata.csv"
  if (!file.exists(alldata_path)) {
    GET("https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/asrh/cc-est2019-alldata.csv",
        add_headers(`User-Agent` = "Mozilla/5.0"), write_disk(alldata_path, overwrite = TRUE), timeout(300))
  }
  d19 <- read_csv(alldata_path, show_col_types = FALSE, col_types = cols(.default = "c")) %>%
    filter(SUMLEV == "050", YEAR == "12", AGEGRP == "0") %>%
    transmute(fips = paste0(sprintf("%02s", STATE), sprintf("%03s", COUNTY)),
      total_2019 = as.numeric(TOT_POP),
      white = as.numeric(WA_MALE) + as.numeric(WA_FEMALE), black = as.numeric(BA_MALE) + as.numeric(BA_FEMALE),
      aian = as.numeric(IA_MALE) + as.numeric(IA_FEMALE), asian = as.numeric(AA_MALE) + as.numeric(AA_FEMALE),
      nhpi = as.numeric(`NA_MALE`) + as.numeric(`NA_FEMALE`), tom = as.numeric(TOM_MALE) + as.numeric(TOM_FEMALE),
      hisp = as.numeric(H_MALE) + as.numeric(H_FEMALE))

  race_2019 <- d19 %>% transmute(fips, total_2019,
    white_share_2019 = 100*white/total_2019, black_share_2019 = 100*black/total_2019,
    aian_share_2019 = 100*aian/total_2019, asian_share_2019 = 100*asian/total_2019,
    nhpi_share_2019 = 100*nhpi/total_2019, tom_share_2019 = 100*tom/total_2019,
    hispanic_share_2019 = 100*hisp/total_2019)
  saveRDS(race_2019, race_2019_path)
}
cat("2019 race table:", nrow(race_2019), "counties (full PEP coverage)\n")
```

Race studied bucket: `white_share` dropped as the reference category (shares
sum to 100 by construction, so all six can't enter a regression together).
AIAN + NHPI + TOM collapsed into a single `other_race_share` -- individually
these are frequently near-zero, noisy county-level shares. Final bucket:
`black_share`, `hispanic_share`, `asian_share`, `other_race_share`.

## 2. New predictor: EDUCATION (ACS)

2005-2007 3-year for the 2008 crisis, 2015-2019 5-year for COVID. Bachelor's
degree or higher, population 25+.

- **2005-2007 (acs3)**: table **B15002** ("Sex by Educational Attainment"),
  not B15003 -- confirmed live, B15003 returns HTTP 404 against the 2007
  acs3 endpoint (it didn't exist as a table until the 2008 vintage; same
  reason the existing 1-year 2007 pull in `pipeline/13_acs_education.R`
  uses B15002). Bachelor's-plus = male cells 015-018 + female cells 032-035.
  **ACS 3-year publishes only for geographies with population >= 20,000** --
  counties below that threshold are absent from the API response entirely
  (not an error, not a 0) -- this is the sample-size-limiting step for this
  entire Option.
- **2015-2019 (acs5)**: table **B15003** (confirmed present, same cell
  layout as the existing 1-year 2019 pull: total = cell 001, bachelor's-plus
  = cells 022-025). ACS 5-year has full county coverage, no threshold.

```{r education-2007, cache=TRUE}
edu_2007_path <- "raw_data/optionc_education_2007.rds"
if (file.exists(edu_2007_path)) {
  edu_2007 <- readRDS(edu_2007_path)
} else {
  acs_get <- function(base, vars, geo) {
    url <- paste0(base, "?get=NAME,", paste(vars, collapse = ","), "&", geo, "&key=", census_key)
    raw <- fromJSON(url)
    d <- as_tibble(raw[-1, , drop = FALSE], .name_repair = "minimal")
    names(d) <- raw[1, ]
    d
  }
  male_ba <- paste0("B15002_0", 15:18, "E"); female_ba <- paste0("B15002_0", 32:35, "E")
  vars07 <- c("B15002_001E", male_ba, female_ba)
  edu07_raw <- acs_get("https://api.census.gov/data/2007/acs/acs3", vars07, "for=county:*&in=state:*")

  edu_2007 <- edu07_raw %>%
    mutate(across(all_of(vars07), as.numeric)) %>%
    rowwise() %>%
    mutate(ba_plus = sum(c_across(all_of(c(male_ba, female_ba))), na.rm = TRUE)) %>%
    ungroup() %>%
    filter(state != "72") %>%
    transmute(fips = paste0(state, county), bachelors_share_2007 = 100 * ba_plus / B15002_001E)
  saveRDS(edu_2007, edu_2007_path)
}
cat("2007 education (ACS3 2005-2007, 20k+ threshold, post-PR):", nrow(edu_2007), "counties\n")
```

```{r education-2019, cache=TRUE}
edu_2019_path <- "raw_data/optionc_education_2019.rds"
if (file.exists(edu_2019_path)) {
  edu_2019 <- readRDS(edu_2019_path)
} else {
  acs_get <- function(base, vars, geo) {
    url <- paste0(base, "?get=NAME,", paste(vars, collapse = ","), "&", geo, "&key=", census_key)
    raw <- fromJSON(url)
    d <- as_tibble(raw[-1, , drop = FALSE], .name_repair = "minimal")
    names(d) <- raw[1, ]
    d
  }
  ba_2019 <- paste0("B15003_0", 22:25, "E")
  vars19 <- c("B15003_001E", ba_2019)
  edu19_raw <- acs_get("https://api.census.gov/data/2019/acs/acs5", vars19, "for=county:*&in=state:*")

  edu_2019 <- edu19_raw %>%
    mutate(across(all_of(vars19), as.numeric)) %>%
    rowwise() %>%
    mutate(ba_plus = sum(c_across(all_of(ba_2019)), na.rm = TRUE)) %>%
    ungroup() %>%
    filter(state != "72") %>%
    transmute(fips = paste0(state, county), bachelors_share_2019 = 100 * ba_plus / B15003_001E)
  saveRDS(edu_2019, edu_2019_path)
}
cat("2019 education (ACS5 2015-2019, full coverage, post-PR):", nrow(edu_2019), "counties\n")
```

## 3. New predictor: INDUSTRY MIX (ACS table C24030, resident-based)

**Resident-based, not establishment-based.** This is "sex by industry for the
civilian employed population 16+, by place of **residence**" -- Option B's
industry variable (`pipeline/20_cbp_industry_mix.R`) used County Business
Patterns instead, which counts jobs by **establishment location**. A county
that's mostly bedroom-community (workers commute out) looks very different
under the two definitions -- this matters for interpretation, not just
sourcing.

**Sector mapping verified, not assumed.** Checked both vintages' variable
lists against the Census API's own `/groups/C24030.json` before building any
share:

```{r industry-mapping-check, cache=TRUE}
get_group_vars <- function(url) {
  raw <- fromJSON(url)
  vars <- raw$variables
  tibble(code = names(vars), label = sapply(vars, function(v) v$label)) %>%
    filter(grepl("E$", code), !grepl("EA$", code)) %>% arrange(code)
}
v2007 <- get_group_vars("https://api.census.gov/data/2007/acs/acs3/groups/C24030.json")
v2019 <- get_group_vars("https://api.census.gov/data/2019/acs/acs5/groups/C24030.json")
cat("2007 acs3 C24030 variable count:", nrow(v2007), " | 2019 acs5:", nrow(v2019), "\n")
cat("Same variable codes present in both vintages:", setequal(v2007$code, v2019$code), "\n")
```

**Result: identical variable-code structure in both vintages** (same 56
codes, code-for-code) -- code `001E` = total, `002E`-`028E` = Male (27
cells), `029E`-`055E` = Female (27 cells, exact +27 offset of the male
block). Label wording shifts slightly between vintages (e.g. 2007's "Mining"
becomes 2019's "Mining, quarrying, and oil and gas extraction" -- a
terminology update, not a category change), but the category composition
and cell numbering are unchanged. **No sector failed to map** -- unlike CBP's
NAICS2002-vs-NAICS2017 vintage break (Option B's problem), C24030 is
designed as a stable industry-supersector table across ACS vintages.

Collapsed to **13 MECE top-level sectors** (the direct children of the
Male/Female total, skipping the more-detailed nested sub-splits so shares
sum cleanly to 100 with no double-counting):

```{r industry-sector-lookup}
sector_lookup <- tibble(
  sector = c("agriculture_mining","construction","manufacturing","wholesale_trade","retail_trade",
             "transportation_utilities","information","finance_real_estate","professional_admin",
             "education_health","arts_accommodation","other_services","public_administration"),
  male_code = c(3,6,7,8,9,10,13,14,17,21,24,27,28),
  female_code = c(3,6,7,8,9,10,13,14,17,21,24,27,28) + 27
)
sector_lookup
```

```{r industry-2007, cache=TRUE}
ind_2007_path <- "raw_data/optionc_industry_2007.rds"
if (file.exists(ind_2007_path)) {
  ind_2007 <- readRDS(ind_2007_path)
} else {
  acs_get <- function(base, vars, geo) {
    url <- paste0(base, "?get=NAME,", paste(vars, collapse = ","), "&", geo, "&key=", census_key)
    raw <- fromJSON(url)
    d <- as_tibble(raw[-1, , drop = FALSE], .name_repair = "minimal")
    names(d) <- raw[1, ]
    d
  }
  all_codes <- c(1, sector_lookup$male_code, sector_lookup$female_code)
  vars <- paste0("C24030_", sprintf("%03d", all_codes), "E")
  ind07_raw <- acs_get("https://api.census.gov/data/2007/acs/acs3", vars, "for=county:*&in=state:*")

  build_shares <- function(raw, suffix) {
    raw <- raw %>% mutate(across(all_of(vars), as.numeric)) %>% filter(state != "72")
    out <- tibble(fips = paste0(raw$state, raw$county))
    total <- raw[["C24030_001E"]]
    for (i in seq_len(nrow(sector_lookup))) {
      mcol <- paste0("C24030_", sprintf("%03d", sector_lookup$male_code[i]), "E")
      fcol <- paste0("C24030_", sprintf("%03d", sector_lookup$female_code[i]), "E")
      out[[paste0(sector_lookup$sector[i], "_share_", suffix)]] <- 100 * (raw[[mcol]] + raw[[fcol]]) / total
    }
    out
  }
  ind_2007 <- build_shares(ind07_raw, "2007")
  saveRDS(ind_2007, ind_2007_path)
}
cat("2007 industry (ACS3 2005-2007, 20k+ threshold):", nrow(ind_2007), "counties\n")
```

```{r industry-2019, cache=TRUE}
ind_2019_path <- "raw_data/optionc_industry_2019.rds"
if (file.exists(ind_2019_path)) {
  ind_2019 <- readRDS(ind_2019_path)
} else {
  acs_get <- function(base, vars, geo) {
    url <- paste0(base, "?get=NAME,", paste(vars, collapse = ","), "&", geo, "&key=", census_key)
    raw <- fromJSON(url)
    d <- as_tibble(raw[-1, , drop = FALSE], .name_repair = "minimal")
    names(d) <- raw[1, ]
    d
  }
  all_codes <- c(1, sector_lookup$male_code, sector_lookup$female_code)
  vars <- paste0("C24030_", sprintf("%03d", all_codes), "E")
  ind19_raw <- acs_get("https://api.census.gov/data/2019/acs/acs5", vars, "for=county:*&in=state:*")

  build_shares <- function(raw, suffix) {
    raw <- raw %>% mutate(across(all_of(vars), as.numeric)) %>% filter(state != "72")
    out <- tibble(fips = paste0(raw$state, raw$county))
    total <- raw[["C24030_001E"]]
    for (i in seq_len(nrow(sector_lookup))) {
      mcol <- paste0("C24030_", sprintf("%03d", sector_lookup$male_code[i]), "E")
      fcol <- paste0("C24030_", sprintf("%03d", sector_lookup$female_code[i]), "E")
      out[[paste0(sector_lookup$sector[i], "_share_", suffix)]] <- 100 * (raw[[mcol]] + raw[[fcol]]) / total
    }
    out
  }
  ind_2019 <- build_shares(ind19_raw, "2019")
  saveRDS(ind_2019, ind_2019_path)
}
cat("2019 industry (ACS5 2015-2019, full coverage):", nrow(ind_2019), "counties\n")
```

Industry studied bucket: 13 shares sum to 100, so one must drop as reference
-- **`public_administration_share` dropped** (least tied to private-sector
business-cycle dynamics, a reasonable stable baseline). Final bucket: the
other 12 sector shares.

## 4. Assemble spine_data_optionc.csv

Read-only on `spine_data.csv` (Option A's file) and on every Option A
artifact. Output is a brand-new file, `spine_data_optionc.csv`, in the
project root.

```{r assemble-spine}
spine_optionc_path <- "spine_data_optionc.csv"
if (file.exists(spine_optionc_path)) {
  spine_optionc <- read_csv(spine_optionc_path, show_col_types = FALSE)
} else {
  county <- read_csv("spine_data.csv", show_col_types = FALSE)  # Option A's file -- read-only

  race_2007_bucket <- race_2007 %>% transmute(fips,
    black_share_2007, hispanic_share_2007, asian_share_2007,
    other_race_share_2007 = aian_share_2007 + nhpi_share_2007 + tom_share_2007)
  race_2019_bucket <- race_2019 %>% transmute(fips,
    black_share_2019, hispanic_share_2019, asian_share_2019,
    other_race_share_2019 = aian_share_2019 + nhpi_share_2019 + tom_share_2019)

  # NB: select(-col), not select(fips, -col) -- the latter keeps ONLY fips
  ind_2007_bucket <- ind_2007 %>% select(-public_administration_share_2007)
  ind_2019_bucket <- ind_2019 %>% select(-public_administration_share_2019)

  spine_optionc <- county %>%
    left_join(race_2007_bucket, by = "fips") %>%
    left_join(race_2019_bucket, by = "fips") %>%
    left_join(edu_2007, by = "fips") %>%
    left_join(edu_2019, by = "fips") %>%
    left_join(ind_2007_bucket, by = "fips") %>%
    left_join(ind_2019_bucket, by = "fips")

  write_csv(spine_optionc, spine_optionc_path)
}
cat("spine_data_optionc.csv:", nrow(spine_optionc), "rows,", ncol(spine_optionc), "cols\n")
```

## 5. Sample: intersection of complete cases across both crises

```{r sample-intersection}
# public_administration was dropped as the reference category when the spine
# was assembled (Section 4) -- only the other 12 sectors actually exist as
# columns in spine_optionc, so this list must match that, not sector_lookup's
# full 13.
industry_sectors <- setdiff(sector_lookup$sector, "public_administration")

crisis_cols <- list(
  "2008" = c(outcome = "recovery_ratio_2008",
             proprietor_share = "proprietor_share_2007", migration_rate = "net_migration_rate_2007",
             unemployment_rate = "pre_unemployment_2007", working_age_share = "working_age_share_2007",
             income = "median_hh_income_2007", poverty_rate = "poverty_rate_2007",
             population = "total_population_2007", density = "population_density_2007",
             rucc = "rucc_2003_2007", pop_trend = "pop_trend_2007",
             black_share = "black_share_2007", hispanic_share = "hispanic_share_2007",
             asian_share = "asian_share_2007", other_race_share = "other_race_share_2007",
             bachelors_share = "bachelors_share_2007"),
  "covid" = c(outcome = "recovery_ratio_covid",
              proprietor_share = "proprietor_share_2019", migration_rate = "net_migration_rate_2019",
              unemployment_rate = "pre_unemployment_2019", working_age_share = "working_age_share_2019",
              income = "median_hh_income_2019", poverty_rate = "poverty_rate_2019",
              population = "total_population_2019", density = "population_density_2019",
              rucc = "rucc_2013_2019", pop_trend = "pop_trend_2019",
              black_share = "black_share_2019", hispanic_share = "hispanic_share_2019",
              asian_share = "asian_share_2019", other_race_share = "other_race_share_2019",
              bachelors_share = "bachelors_share_2019")
)

req_cols_for <- function(crisis) {
  suffix <- if (crisis == "2008") "2007" else "2019"
  c(unname(crisis_cols[[crisis]]), paste0(industry_sectors, "_share_", suffix))
}

# NA count per required column -- identifies which predictor drives sample loss
na_report_2008 <- sapply(req_cols_for("2008"), function(cl) sum(is.na(spine_optionc[[cl]])))
na_report_covid <- sapply(req_cols_for("covid"), function(cl) sum(is.na(spine_optionc[[cl]])))
cat("=== 2008: top NA-count columns ===\n"); print(sort(na_report_2008, decreasing = TRUE)[1:6])
cat("\n=== COVID: top NA-count columns ===\n"); print(sort(na_report_covid, decreasing = TRUE)[1:6])

complete_fips <- function(crisis) {
  cols <- req_cols_for(crisis)
  spine_optionc %>% filter(if_all(all_of(cols), ~ !is.na(.))) %>% pull(fips)
}
fips_2008 <- complete_fips("2008")
fips_covid <- complete_fips("covid")
fips_final <- intersect(fips_2008, fips_covid)

cat("\nComplete cases, 2008-side predictors:", length(fips_2008), "\n")
cat("Complete cases, COVID-side predictors:", length(fips_covid), "\n")
cat("INTERSECTION (final Option C sample, used for both crises):", length(fips_final), "\n")
cat("2008-complete counties dropped by intersection:", length(setdiff(fips_2008, fips_final)), "\n")
cat("COVID-complete counties dropped by intersection:", length(setdiff(fips_covid, fips_final)), "\n")

saveRDS(fips_final, "raw_data/optionc_final_fips.rds")
```

**The ACS 2005-2007 3-year 20,000+ population threshold is the dominant
driver of sample loss** -- `bachelors_share_2007` and all 12
`*_share_2007` industry columns are simultaneously NA for the same ~1,400+
counties (every county below the threshold, at once), dwarfing the
pre-existing Option A missingness (~90-150 per column) in the original 4
predictors/controls. The COVID side barely constrains anything (ACS5 has
full coverage) -- every 2008-complete county is also COVID-complete, so the
intersection equals the 2008-side count exactly.

## 6. Diagnostics on the 3 new predictor buckets

```{r diagnostics-distributions}
d_diag <- spine_optionc %>% filter(fips %in% fips_final)
new_2007 <- c("black_share_2007","hispanic_share_2007","asian_share_2007","other_race_share_2007",
  "bachelors_share_2007", paste0(industry_sectors[industry_sectors != "public_administration"], "_share_2007"))

dist_table <- t(sapply(new_2007, function(v) {
  x <- d_diag[[v]]
  c(mean = mean(x, na.rm=TRUE), sd = sd(x, na.rm=TRUE), min = min(x, na.rm=TRUE),
    median = median(x, na.rm=TRUE), max = max(x, na.rm=TRUE), n_na = sum(is.na(x)))
}))
round(dist_table, 2)
```

```{r diagnostics-correlations}
d_diag <- d_diag %>% mutate(log_population_2007 = log(total_population_2007),
                             log_density_2007 = log(population_density_2007))
existing_cols <- c("proprietor_share_2007","net_migration_rate_2007","pre_unemployment_2007",
  "working_age_share_2007","median_hh_income_2007","poverty_rate_2007","log_population_2007",
  "log_density_2007","pop_trend_2007")

cor_mat <- cor(d_diag[new_2007], d_diag[existing_cols], use = "complete.obs")
flagged <- which(abs(cor_mat) > 0.6, arr.ind = TRUE)
cat("=== |r| > 0.6 (new predictor vs. existing predictor/control) ===\n")
for (i in seq_len(nrow(flagged))) {
  r <- flagged[i,1]; cc <- flagged[i,2]
  cat(sprintf("%-32s vs %-25s : r = %.3f\n", new_2007[r], existing_cols[cc], cor_mat[r,cc]))
}
```

**Flagged: `bachelors_share` vs `income` (r=0.68)** -- exactly the pairing
the task anticipated. `professional_admin_share` also correlates with income
(0.65) and log(population) (0.64) -- professional/administrative jobs
concentrate in wealthier, denser counties, unsurprisingly. Nothing else
clears 0.6; the rest of the industry/race shares are only weakly correlated
with the existing predictor/control set.

## 7. Prepare crisis-specific data (same ~1,766-county sample, both crises)

```{r prepare-crisis-data}
prepare_crisis_data_optionc <- function(raw, crisis, fips_keep) {
  cols <- crisis_cols[[crisis]]
  suffix <- if (crisis == "2008") "2007" else "2019"
  industry_src_cols <- paste0(industry_sectors, "_share_", suffix)

  base <- raw %>%
    filter(fips %in% fips_keep) %>%
    transmute(
      fips, county_name,
      outcome_raw = .data[[cols[["outcome"]]]],
      proprietor_share = .data[[cols[["proprietor_share"]]]],
      migration_rate = .data[[cols[["migration_rate"]]]],
      unemployment_rate = .data[[cols[["unemployment_rate"]]]],
      working_age_share = .data[[cols[["working_age_share"]]]],
      income = .data[[cols[["income"]]]], poverty_rate = .data[[cols[["poverty_rate"]]]],
      population = .data[[cols[["population"]]]], density = .data[[cols[["density"]]]],
      rucc_raw = .data[[cols[["rucc"]]]], pop_trend = .data[[cols[["pop_trend"]]]],
      black_share = .data[[cols[["black_share"]]]], hispanic_share = .data[[cols[["hispanic_share"]]]],
      asian_share = .data[[cols[["asian_share"]]]], other_race_share = .data[[cols[["other_race_share"]]]],
      bachelors_share = .data[[cols[["bachelors_share"]]]]
    )

  industry_raw <- raw %>% filter(fips %in% fips_keep) %>% select(fips, all_of(industry_src_cols))
  names(industry_raw) <- c("fips", industry_sectors)

  df <- base %>% left_join(industry_raw, by = "fips") %>%
    mutate(
      log_outcome = log(outcome_raw), log_population = log(population), log_density = log(density),
      rucc = relevel(factor(rucc_raw), ref = "1"),
      proprietor_share_z = as.numeric(scale(proprietor_share)),
      migration_rate_z = as.numeric(scale(migration_rate)),
      unemployment_rate_z = as.numeric(scale(unemployment_rate)),
      working_age_share_z = as.numeric(scale(working_age_share)),
      black_share_z = as.numeric(scale(black_share)), hispanic_share_z = as.numeric(scale(hispanic_share)),
      asian_share_z = as.numeric(scale(asian_share)), other_race_share_z = as.numeric(scale(other_race_share)),
      bachelors_share_z = as.numeric(scale(bachelors_share))
    )
  for (s in industry_sectors) df[[paste0(s, "_z")]] <- as.numeric(scale(df[[s]]))
  df
}

df_2008 <- prepare_crisis_data_optionc(spine_optionc, "2008", fips_final)
df_covid <- prepare_crisis_data_optionc(spine_optionc, "covid", fips_final)
cat("df_2008:", nrow(df_2008), "| df_covid:", nrow(df_covid),
    "| identical fips sets:", setequal(df_2008$fips, df_covid$fips), "\n")
```

## 8. Model set -- FULL (21 studied predictors) and 4-PREDICTOR-ONLY, same sample

Two predictor sets, both run through the complete Option A model suite (OLS
HC1, LASSO with `penalty.factor=0` on controls, relaxed LASSO, RF with/
without wealth, wealth-band OLS), on the **identical** ~1,766-county sample
-- the 4-predictor-only rerun isolates whether any change from Option A is
driven by the new predictors or just by the smaller sample.

```{r run-crisis-models-optionc}
studied_4 <- c("proprietor_share_z","migration_rate_z","unemployment_rate_z","working_age_share_z")
studied_new <- c("black_share_z","hispanic_share_z","asian_share_z","other_race_share_z",
                  "bachelors_share_z", paste0(industry_sectors, "_z"))
studied_full <- c(studied_4, studied_new)

controls <- c("income","poverty_rate","log_population","log_density","rucc","pop_trend")
controls_no_wealth <- setdiff(controls, c("income","poverty_rate"))

run_crisis_models_optionc <- function(df, studied, seed = 20260729) {
  ols_formula <- as.formula(paste("log_outcome ~", paste(c(studied, controls), collapse = " + ")))
  ols <- lm_robust(ols_formula, data = df, se_type = "HC1")

  X <- model.matrix(as.formula(paste("~", paste(c(studied, controls), collapse = " + "))), data = df)[, -1]
  y <- df$log_outcome
  penalty <- ifelse(colnames(X) %in% studied, 1, 0)
  set.seed(seed)
  cvfit <- cv.glmnet(X, y, alpha = 1, nfolds = 10, standardize = TRUE, penalty.factor = penalty)
  set.seed(seed)
  relaxed <- cv.glmnet(X, y, alpha = 1, nfolds = 10, standardize = TRUE, penalty.factor = penalty, relax = TRUE)

  rf_formula_a <- as.formula(paste("log_outcome ~", paste(c(studied, controls), collapse = " + ")))
  set.seed(seed)
  rf_a <- randomForest(rf_formula_a, data = df, ntree = 500, importance = FALSE)

  rf_formula_b <- as.formula(paste("log_outcome ~", paste(c(studied, controls_no_wealth), collapse = " + ")))
  set.seed(seed)
  rf_b <- randomForest(rf_formula_b, data = df, ntree = 500, importance = TRUE)

  df_banded <- df %>% mutate(income_band = cut(income, breaks = quantile(income, c(0, 1/3, 2/3, 1)),
                                                include.lowest = TRUE, labels = c("lower","middle","higher")))
  band_fits <- setNames(lapply(c("lower","middle","higher"), function(b) {
    lm_robust(ols_formula, data = df_banded %>% filter(income_band == b), se_type = "HC1")
  }), c("lower","middle","higher"))

  list(n = nrow(df), studied = studied, ols = ols, cvfit = cvfit, relaxed = relaxed,
       coef_min = as.matrix(coef(cvfit, s = "lambda.min"))[, 1],
       coef_1se = as.matrix(coef(cvfit, s = "lambda.1se"))[, 1],
       rf_a = rf_a, rf_b = rf_b, band_fits = band_fits, df_banded = df_banded)
}

model_cache <- "raw_data/optionc_model_results.rds"
if (file.exists(model_cache)) {
  model_results <- readRDS(model_cache)
} else {
  message("Fitting FULL (21-predictor) models...")
  res_2008_full <- run_crisis_models_optionc(df_2008, studied_full)
  res_covid_full <- run_crisis_models_optionc(df_covid, studied_full)
  message("Fitting 4-predictor-only models (same sample)...")
  res_2008_4only <- run_crisis_models_optionc(df_2008, studied_4)
  res_covid_4only <- run_crisis_models_optionc(df_covid, studied_4)
  model_results <- list(full = list("2008" = res_2008_full, "covid" = res_covid_full),
                         four_only = list("2008" = res_2008_4only, "covid" = res_covid_4only),
                         studied_4 = studied_4, studied_full = studied_full)
  saveRDS(model_results, model_cache)
}
```

## 9. Results: education, industry, race -- OLS (full model)

```{r results-ols-new-predictors}
edu_ind_vars <- c("bachelors_share_z", paste0(industry_sectors[industry_sectors != "public_administration"], "_z"))
race_vars <- c("black_share_z","hispanic_share_z","asian_share_z","other_race_share_z")

show_ols <- function(fit, vars) {
  s <- summary(fit)$coefficients
  round(s[rownames(s) %in% vars, c("Estimate","Std. Error","Pr(>|t|)"), drop = FALSE], 4)
}
cat("--- 2008: education, industry, race ---\n")
show_ols(model_results$full[["2008"]]$ols, c(edu_ind_vars, race_vars))
cat("\n--- COVID: education, industry, race ---\n")
show_ols(model_results$full[["covid"]]$ols, c(edu_ind_vars, race_vars))
```

## 10. Results: original 4 predictors -- Option A vs. Option C (4-only) vs. Option C (full)

```{r results-original-4-comparison}
optionA <- readRDS("raw_data/modeling_andrew_results.rds")  # Option A's cache -- read-only

get_ols4 <- function(fit) {
  s <- summary(fit)$coefficients[studied_4, c("Estimate","Std. Error","Pr(>|t|)"), drop = FALSE]
  round(s, 5)
}
cat("=== OPTION A (original spine, n=", optionA[["2008"]]$n, "/", optionA[["covid"]]$n, ") ===\n")
cat("-- 2008 --\n"); print(get_ols4(optionA[["2008"]]$ols))
cat("-- COVID --\n"); print(get_ols4(optionA[["covid"]]$ols))

cat("\n=== OPTION C, 4-predictor-only, same n=", model_results$four_only[["2008"]]$n, "sample ===\n")
cat("-- 2008 --\n"); print(get_ols4(model_results$four_only[["2008"]]$ols))
cat("-- COVID --\n"); print(get_ols4(model_results$four_only[["covid"]]$ols))

cat("\n=== OPTION C, FULL 21-predictor model, same sample ===\n")
cat("-- 2008 --\n"); print(get_ols4(model_results$full[["2008"]]$ols))
cat("-- COVID --\n"); print(get_ols4(model_results$full[["covid"]]$ols))
```

## 11. Results: wealth bands (full model)

```{r results-wealth-bands}
band_n <- function(r) sapply(r$band_fits, function(f) f$nobs)
cat("Band n, 2008:", band_n(model_results$full[["2008"]]), " | COVID:", band_n(model_results$full[["covid"]]), "\n\n")

show_bands <- function(r, label) {
  cat("---", label, "---\n")
  for (b in c("lower","middle","higher")) {
    s <- summary(r$band_fits[[b]])$coefficients[studied_4, c("Estimate","Pr(>|t|)"), drop = FALSE]
    cat(" band:", b, "n =", r$band_fits[[b]]$nobs, "\n")
    print(round(s, 4))
  }
}
show_bands(model_results$full[["2008"]], "2008 FULL -- wealth bands")
show_bands(model_results$full[["covid"]], "COVID FULL -- wealth bands")
```

## 12. LASSO survival check (education/industry/race, full model, lambda.min)

```{r results-lasso-survival}
for (crisis in c("2008","covid")) {
  cat("===", crisis, "===\n")
  cm <- model_results$full[[crisis]]$coef_min[c(edu_ind_vars, race_vars)]
  cat("Nonzero:\n"); print(round(cm[cm != 0], 5))
  cat("Zeroed out:", paste(names(cm[cm == 0]), collapse = ", "), "\n\n")
}
```

# Verification-pass fixes

Two problems surfaced in a verification pass before A/C synthesis: (1) the 4
original predictors were standardized on Option C's own (smaller) sample,
not Option A's, so standardized coefficients from A and C were not on the
same scale; (2) RUCC 8 and 9 had only 12-22 counties each in the Option C
sample (from the ACS 20k+ threshold gutting the most-rural categories).
Sections 13+ below fix both, add the one missing correlation number, and run
a cross-crisis prediction test. Nothing above is modified -- these are new,
additive sections. New fits below save to `raw_data/optionc_*_v2.rds` and
`raw_data/crosscrisis_*.rds` only; Option A's four protected `.rds` files are
read-only everywhere they're referenced.

## 13. Fix 1 + Fix 2: rescale to Option A, collapse RUCC 8/9

```{r fix-prep-data}
# Option A's full-sample mean/SD for the 4 original predictors, per crisis --
# the scale the 4 original predictors will be standardized on everywhere
# below, so Option A and Option C coefficients are directly comparable.
raw_vars <- c(proprietor_share_z = "proprietor_share", migration_rate_z = "migration_rate",
              unemployment_rate_z = "unemployment_rate", working_age_share_z = "working_age_share")
optionA_scale <- list(
  "2008" = sapply(raw_vars, function(v) c(mean = mean(optionA[["2008"]]$df[[v]]), sd = sd(optionA[["2008"]]$df[[v]]))),
  "covid" = sapply(raw_vars, function(v) c(mean = mean(optionA[["covid"]]$df[[v]]), sd = sd(optionA[["covid"]]$df[[v]])))
)

prepare_crisis_data_v2 <- function(raw, crisis, fips_keep) {
  cols <- crisis_cols[[crisis]]
  suffix <- if (crisis == "2008") "2007" else "2019"
  industry_src_cols <- paste0(industry_sectors, "_share_", suffix)

  base <- raw %>%
    filter(fips %in% fips_keep) %>%
    transmute(
      fips, county_name,
      outcome_raw = .data[[cols[["outcome"]]]],
      proprietor_share = .data[[cols[["proprietor_share"]]]],
      migration_rate = .data[[cols[["migration_rate"]]]],
      unemployment_rate = .data[[cols[["unemployment_rate"]]]],
      working_age_share = .data[[cols[["working_age_share"]]]],
      income = .data[[cols[["income"]]]], poverty_rate = .data[[cols[["poverty_rate"]]]],
      population = .data[[cols[["population"]]]], density = .data[[cols[["density"]]]],
      rucc_raw = .data[[cols[["rucc"]]]], pop_trend = .data[[cols[["pop_trend"]]]],
      black_share = .data[[cols[["black_share"]]]], hispanic_share = .data[[cols[["hispanic_share"]]]],
      asian_share = .data[[cols[["asian_share"]]]], other_race_share = .data[[cols[["other_race_share"]]]],
      bachelors_share = .data[[cols[["bachelors_share"]]]]
    )

  industry_raw <- raw %>% filter(fips %in% fips_keep) %>% select(fips, all_of(industry_src_cols))
  names(industry_raw) <- c("fips", industry_sectors)

  scaleA <- optionA_scale[[crisis]]
  df <- base %>% left_join(industry_raw, by = "fips") %>%
    mutate(
      log_outcome = log(outcome_raw), log_population = log(population), log_density = log(density),
      # Fix 2: RUCC 8 and 9 collapsed into one level (12/13 and 22/13 counties
      # respectively -- too thin on their own)
      rucc_collapsed = ifelse(rucc_raw %in% c(8, 9), "8_9", as.character(rucc_raw)),
      rucc = relevel(factor(rucc_collapsed), ref = "1"),
      # Fix 1: 4 original predictors standardized on OPTION A's full-sample
      # mean/SD, not Option C's own -- makes coefficients directly comparable
      proprietor_share_z = (proprietor_share - scaleA["mean","proprietor_share_z"]) / scaleA["sd","proprietor_share_z"],
      migration_rate_z = (migration_rate - scaleA["mean","migration_rate_z"]) / scaleA["sd","migration_rate_z"],
      unemployment_rate_z = (unemployment_rate - scaleA["mean","unemployment_rate_z"]) / scaleA["sd","unemployment_rate_z"],
      working_age_share_z = (working_age_share - scaleA["mean","working_age_share_z"]) / scaleA["sd","working_age_share_z"],
      # new predictors have no Option A equivalent -- standardized on Option C's own sample
      black_share_z = as.numeric(scale(black_share)), hispanic_share_z = as.numeric(scale(hispanic_share)),
      asian_share_z = as.numeric(scale(asian_share)), other_race_share_z = as.numeric(scale(other_race_share)),
      bachelors_share_z = as.numeric(scale(bachelors_share))
    )
  for (s in industry_sectors) df[[paste0(s, "_z")]] <- as.numeric(scale(df[[s]]))
  df
}

prep_v2_cache <- "raw_data/optionc_prepared_data_v2.rds"
if (file.exists(prep_v2_cache)) {
  prep_v2 <- readRDS(prep_v2_cache)
} else {
  df_2008_v2 <- prepare_crisis_data_v2(spine_optionc, "2008", fips_final)
  df_covid_v2 <- prepare_crisis_data_v2(spine_optionc, "covid", fips_final)
  prep_v2 <- list(df_2008 = df_2008_v2, df_covid = df_covid_v2, industry_sectors = industry_sectors,
                   optionA_scale = optionA_scale)
  saveRDS(prep_v2, prep_v2_cache)
}
cat("RUCC levels, 2008 (post-collapse):\n"); print(table(prep_v2$df_2008$rucc))
cat("RUCC levels, COVID (post-collapse):\n"); print(table(prep_v2$df_covid$rucc))
```

## 14. Refit Option C models (4-only and full) on the fixed data

```{r fix-refit-models}
model_v2_cache <- "raw_data/optionc_model_results_v2.rds"
if (file.exists(model_v2_cache)) {
  model_results_v2 <- readRDS(model_v2_cache)
} else {
  # penalty.factor rebuilt from colnames(model.matrix()) -- automatically
  # picks up the reduced RUCC dummy count (7, not 8) after the collapse
  res_2008_full_v2 <- run_crisis_models_optionc(prep_v2$df_2008, studied_full)
  res_covid_full_v2 <- run_crisis_models_optionc(prep_v2$df_covid, studied_full)
  res_2008_4only_v2 <- run_crisis_models_optionc(prep_v2$df_2008, studied_4)
  res_covid_4only_v2 <- run_crisis_models_optionc(prep_v2$df_covid, studied_4)
  model_results_v2 <- list(full = list("2008" = res_2008_full_v2, "covid" = res_covid_full_v2),
                            four_only = list("2008" = res_2008_4only_v2, "covid" = res_covid_4only_v2))
  saveRDS(model_results_v2, model_v2_cache)
}

X_check <- model.matrix(as.formula(paste("~", paste(c(studied_4, controls), collapse=" + "))), data = prep_v2$df_2008)
cat("RUCC dummy columns post-collapse:", sum(grepl("^rucc", colnames(X_check))), "(was 8, now 7)\n")
```

## 15. Rescaled comparison: Option A vs. Option C (4-only) vs. Option C (full)

```{r fix-comparison-table}
get4 <- function(fit) round(summary(fit)$coefficients[studied_4, c("Estimate","Std. Error","Pr(>|t|)")], 5)

cat("=== OPTION A ===\n")
cat("-- 2008 --\n"); print(get4(optionA[["2008"]]$ols))
cat("-- COVID --\n"); print(get4(optionA[["covid"]]$ols))

cat("\n=== OPTION C v2, 4-only (rescaled to A, RUCC collapsed) ===\n")
cat("-- 2008 --\n"); print(get4(model_results_v2$four_only[["2008"]]$ols))
cat("-- COVID --\n"); print(get4(model_results_v2$four_only[["covid"]]$ols))

cat("\n=== OPTION C v2, FULL (rescaled to A, RUCC collapsed) ===\n")
cat("-- 2008 --\n"); print(get4(model_results_v2$full[["2008"]]$ols))
cat("-- COVID --\n"); print(get4(model_results_v2$full[["covid"]]$ols))
```

## 16. Missing number: education vs. the 4 original predictors

```{r fix-education-correlations}
check_cors <- function(df, label) {
  cat("===", label, "===\n")
  for (v in c("proprietor_share","migration_rate","unemployment_rate","working_age_share")) {
    cat(sprintf("bachelors_share vs %-20s r = %.4f\n", v, cor(df$bachelors_share, df[[v]])))
  }
}
check_cors(prep_v2$df_2008, "2008 (n=1766)")
check_cors(prep_v2$df_covid, "COVID (n=1766)")
```

## 17. VIF on the fixed (RUCC-collapsed) full model

```{r fix-vif}
library(car)
vif_v2_cache <- "raw_data/optionc_vif_v2.rds"
if (file.exists(vif_v2_cache)) {
  vif_v2 <- readRDS(vif_v2_cache)
} else {
  fit_formula_full <- as.formula(paste("log_outcome ~", paste(c(studied_full, controls), collapse = " + ")))
  lm_2008_v2 <- lm(fit_formula_full, data = prep_v2$df_2008)
  lm_covid_v2 <- lm(fit_formula_full, data = prep_v2$df_covid)
  get_vif_val <- function(v) if (is.matrix(v)) v[,3]^2 else v
  vif_v2 <- list(vif_2008 = get_vif_val(vif(lm_2008_v2)), vif_covid = get_vif_val(vif(lm_covid_v2)))
  saveRDS(vif_v2, vif_v2_cache)
}
cat("2008 VIF > 5:\n"); print(round(vif_v2$vif_2008[vif_v2$vif_2008 > 5], 2))
cat("COVID VIF > 5:\n"); print(round(vif_v2$vif_covid[vif_v2$vif_covid > 5], 2))
```

## 18. Cross-crisis test: train on one crisis, predict the other

Random forest, 4 original predictors + 6 controls (raw units, not
standardized -- RF splits are scale-invariant, and raw units sidestep any
cross-crisis rescaling mismatch), fit on Option A's own data restricted to
the 3,012-county both-crises intersection (fair comparison: identical
counties in both directions). New fits only -- saved to
`raw_data/crosscrisis_rf_results.rds`, Option A's cached fits untouched.

```{r crosscrisis-test}
crosscrisis_cache <- "raw_data/crosscrisis_rf_results.rds"
if (file.exists(crosscrisis_cache)) {
  cc <- readRDS(crosscrisis_cache)
} else {
  fips_A_both <- intersect(optionA[["2008"]]$df$fips, optionA[["covid"]]$df$fips)
  controls_raw <- c("income","poverty_rate","log_population","log_density","rucc","pop_trend")
  predictors_raw <- c("proprietor_share","migration_rate","unemployment_rate","working_age_share")

  prep_matched <- function(df, fips_keep) {
    df %>% filter(fips %in% fips_keep) %>%
      mutate(rucc = factor(as.character(rucc_raw), levels = as.character(1:9)))
  }
  df_2008_m <- prep_matched(optionA[["2008"]]$df, fips_A_both)
  df_covid_m <- prep_matched(optionA[["covid"]]$df, fips_A_both)

  rf_formula <- as.formula(paste("log_outcome ~", paste(c(predictors_raw, controls_raw), collapse = " + ")))
  set.seed(20260729)
  rf_train2008 <- randomForest(rf_formula, data = df_2008_m, ntree = 500, importance = FALSE)
  set.seed(20260729)
  rf_train_covid <- randomForest(rf_formula, data = df_covid_m, ntree = 500, importance = FALSE)

  r2_rmse <- function(pred, actual) {
    ss_res <- sum((actual - pred)^2); ss_tot <- sum((actual - mean(actual))^2)
    c(R2 = 1 - ss_res/ss_tot, RMSE = sqrt(mean((actual - pred)^2)))
  }

  pred_covid_from_2008 <- predict(rf_train2008, newdata = df_covid_m)
  cross_2008_to_covid <- r2_rmse(pred_covid_from_2008, df_covid_m$log_outcome)
  spearman_2008_to_covid <- cor(pred_covid_from_2008, df_covid_m$log_outcome, method = "spearman")

  pred_covid_oob <- predict(rf_train_covid)  # OOB predictions -- honest same-crisis ceiling
  ceiling_covid <- r2_rmse(pred_covid_oob, df_covid_m$log_outcome)
  spearman_ceiling_covid <- cor(pred_covid_oob, df_covid_m$log_outcome, method = "spearman")

  pred_2008_from_covid <- predict(rf_train_covid, newdata = df_2008_m)
  cross_covid_to_2008 <- r2_rmse(pred_2008_from_covid, df_2008_m$log_outcome)
  spearman_covid_to_2008 <- cor(pred_2008_from_covid, df_2008_m$log_outcome, method = "spearman")

  pred_2008_oob <- predict(rf_train2008)
  ceiling_2008 <- r2_rmse(pred_2008_oob, df_2008_m$log_outcome)
  spearman_ceiling_2008 <- cor(pred_2008_oob, df_2008_m$log_outcome, method = "spearman")

  cc <- list(fips_matched = fips_A_both, df_2008_m = df_2008_m, df_covid_m = df_covid_m,
             rf_train2008 = rf_train2008, rf_train_covid = rf_train_covid,
             pred_covid_from_2008 = pred_covid_from_2008, pred_2008_from_covid = pred_2008_from_covid,
             cross_2008_to_covid = cross_2008_to_covid, ceiling_covid = ceiling_covid,
             cross_covid_to_2008 = cross_covid_to_2008, ceiling_2008 = ceiling_2008,
             spearman_2008_to_covid = spearman_2008_to_covid, spearman_ceiling_covid = spearman_ceiling_covid,
             spearman_covid_to_2008 = spearman_covid_to_2008, spearman_ceiling_2008 = spearman_ceiling_2008)
  saveRDS(cc, crosscrisis_cache)
}

cat("Matched (both-crises) n:", length(cc$fips_matched), "\n\n")
cat("Train 2008 -> Predict COVID:  R2 =", round(cc$cross_2008_to_covid["R2"],4),
    " RMSE =", round(cc$cross_2008_to_covid["RMSE"],4), " Spearman =", round(cc$spearman_2008_to_covid,4), "\n")
cat("Ceiling (COVID->COVID, OOB):  R2 =", round(cc$ceiling_covid["R2"],4),
    " RMSE =", round(cc$ceiling_covid["RMSE"],4), " Spearman =", round(cc$spearman_ceiling_covid,4), "\n\n")
cat("Train COVID -> Predict 2008:  R2 =", round(cc$cross_covid_to_2008["R2"],4),
    " RMSE =", round(cc$cross_covid_to_2008["RMSE"],4), " Spearman =", round(cc$spearman_covid_to_2008,4), "\n")
cat("Ceiling (2008->2008, OOB):    R2 =", round(cc$ceiling_2008["R2"],4),
    " RMSE =", round(cc$ceiling_2008["RMSE"],4), " Spearman =", round(cc$spearman_ceiling_2008,4), "\n")
```

## 19. Figures (figures_optionc/)

```{r fig-setup-optionc}
studied_labels_4 <- c(proprietor_share_z = "Self-employment", migration_rate_z = "Migration rate",
                       unemployment_rate_z = "Unemployment rate", working_age_share_z = "Working-age share")
pal <- c(blue = "#2a78d6", orange = "#eb6834", aqua = "#1baf7a", yellow = "#eda100")
theme_pub <- function(base_size = 14) {
  theme_minimal(base_size = base_size) +
    theme(panel.grid.minor = element_blank(), panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_line(color = "#e1e0d9", linewidth = 0.3),
          strip.text = element_text(face = "bold", size = rel(0.95)), plot.title = element_blank(),
          legend.position = "bottom", axis.title = element_text(size = rel(0.9)))
}
```

### 19a. decomposition.png

```{r fig-decomposition, fig.width=10, fig.height=6}
get_ci_table_4 <- function(fit, crisis_label, method_label) {
  s <- summary(fit)$coefficients[studied_4, ]
  tibble(term = studied_4, estimate = s[,"Estimate"], ci_low = s[,"CI Lower"], ci_high = s[,"CI Upper"],
         crisis = crisis_label, method = method_label)
}
decomp_data <- bind_rows(
  get_ci_table_4(optionA[["2008"]]$ols, "2008", "Option A"),
  get_ci_table_4(optionA[["covid"]]$ols, "COVID", "Option A"),
  get_ci_table_4(model_results_v2$four_only[["2008"]]$ols, "2008", "Option C, 4-only (rescaled)"),
  get_ci_table_4(model_results_v2$four_only[["covid"]]$ols, "COVID", "Option C, 4-only (rescaled)"),
  get_ci_table_4(model_results_v2$full[["2008"]]$ols, "2008", "Option C, full (rescaled)"),
  get_ci_table_4(model_results_v2$full[["covid"]]$ols, "COVID", "Option C, full (rescaled)")
) %>%
  mutate(predictor = factor(studied_labels_4[term], levels = rev(studied_labels_4)),
         method = factor(method, levels = c("Option A", "Option C, 4-only (rescaled)", "Option C, full (rescaled)")))

method_colors <- c("Option A" = unname(pal["blue"]), "Option C, 4-only (rescaled)" = unname(pal["orange"]),
                    "Option C, full (rescaled)" = unname(pal["aqua"]))
dodge_w <- 0.6
p_decomp <- ggplot(decomp_data, aes(x = estimate, y = predictor, color = method, shape = method)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#898781", linewidth = 0.5) +
  geom_errorbar(aes(xmin = ci_low, xmax = ci_high), width = 0.15, linewidth = 0.6,
                position = position_dodge(width = dodge_w)) +
  geom_point(size = 3, position = position_dodge(width = dodge_w)) +
  scale_color_manual(values = method_colors, name = NULL) +
  scale_shape_manual(values = c("Option A" = 16, "Option C, 4-only (rescaled)" = 17, "Option C, full (rescaled)" = 15), name = NULL) +
  facet_wrap(~crisis) +
  labs(x = "OLS coefficient (log recovery ratio per 1 SD, Option A's SD used throughout), 95% CI", y = NULL) +
  theme_pub() +
  guides(color = guide_legend(nrow = 2), shape = guide_legend(nrow = 2))
ggsave("figures_optionc/decomposition.png", p_decomp, width = 10, height = 6, dpi = 300)
p_decomp
```

### 19b. industry_coefficients.png

```{r fig-industry-coefficients, fig.width=10, fig.height=6}
vars_to_plot <- c("bachelors_share_z", paste0(industry_sectors, "_z"))
var_labels_ind <- c(
  bachelors_share_z = "Education (bachelor's+)",
  agriculture_mining_z = "Agriculture / mining", construction_z = "Construction",
  manufacturing_z = "Manufacturing", wholesale_trade_z = "Wholesale trade", retail_trade_z = "Retail trade",
  transportation_utilities_z = "Transportation / utilities", information_z = "Information",
  finance_real_estate_z = "Finance / real estate", professional_admin_z = "Professional / admin",
  education_health_z = "Education & health services", arts_accommodation_z = "Arts / accommodation / food",
  other_services_z = "Other services"
)
get_ci_table_ind <- function(fit, crisis_label, vif_vec) {
  s <- summary(fit)$coefficients[vars_to_plot, ]
  tibble(term = vars_to_plot, estimate = s[,"Estimate"], ci_low = s[,"CI Lower"], ci_high = s[,"CI Upper"],
         crisis = crisis_label, vif = vif_vec[vars_to_plot], high_vif = vif_vec[vars_to_plot] > 5)
}
ind_data <- bind_rows(
  get_ci_table_ind(model_results_v2$full[["2008"]]$ols, "2008", vif_v2$vif_2008),
  get_ci_table_ind(model_results_v2$full[["covid"]]$ols, "COVID", vif_v2$vif_covid)
) %>% mutate(label = var_labels_ind[term])

order_2008 <- ind_data %>% filter(crisis == "2008") %>% arrange(estimate) %>% pull(label)
ind_data <- ind_data %>% mutate(label = factor(label, levels = order_2008))

p_ind <- ggplot(ind_data, aes(x = estimate, y = label)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#898781", linewidth = 0.5) +
  geom_errorbar(aes(xmin = ci_low, xmax = ci_high), width = 0.15, linewidth = 0.6, color = unname(pal["blue"])) +
  geom_point(aes(shape = high_vif), size = 3, color = unname(pal["blue"]), fill = "white", stroke = 1.1) +
  scale_shape_manual(values = c("FALSE" = 16, "TRUE" = 21), name = NULL,
                      labels = c("FALSE" = "VIF ≤ 5", "TRUE" = "VIF > 5 (open circle)")) +
  facet_wrap(~crisis) +
  labs(x = "OLS coefficient (log recovery ratio per 1 SD), 95% CI -- reference sector: public administration",
       y = NULL) +
  theme_pub()
ggsave("figures_optionc/industry_coefficients.png", p_ind, width = 10, height = 6, dpi = 300)
p_ind
```

### 19c. crosscrisis.png

```{r fig-crosscrisis, fig.width=10, fig.height=6}
cross_data <- bind_rows(
  tibble(actual = cc$df_covid_m$log_outcome, predicted = cc$pred_covid_from_2008,
         panel = "Trained on 2008 -> predict COVID"),
  tibble(actual = cc$df_2008_m$log_outcome, predicted = cc$pred_2008_from_covid,
         panel = "Trained on COVID -> predict 2008")
)
rho_labels <- tibble(
  panel = c("Trained on 2008 -> predict COVID", "Trained on COVID -> predict 2008"),
  rho = c(cc$spearman_2008_to_covid, cc$spearman_covid_to_2008)
) %>% mutate(label = sprintf("Spearman rho = %.3f", rho))
range_data <- cross_data %>% group_by(panel) %>%
  summarise(lo = min(c(actual, predicted)), hi = max(c(actual, predicted)), .groups = "drop")
rho_labels <- rho_labels %>% left_join(range_data, by = "panel")

p_cross <- ggplot(cross_data, aes(x = actual, y = predicted)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#898781", linewidth = 0.5) +
  geom_point(color = unname(pal["blue"]), alpha = 0.35, size = 1.6) +
  geom_text(data = rho_labels, aes(x = lo, y = hi, label = label), hjust = 0, vjust = 1.3,
            inherit.aes = FALSE, size = 4.2) +
  facet_wrap(~panel, scales = "free") +
  labs(x = "Actual log(recovery ratio)", y = "Predicted log(recovery ratio) -- RF trained on the other crisis") +
  theme_pub()
ggsave("figures_optionc/crosscrisis.png", p_cross, width = 10, height = 6, dpi = 300)
p_cross
```
